---
title: "Environmental indicator analysis-Tilefish"
output:  
  html_document:
    code_fold: hide
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, wrangling}

# Loading packages and data
library(ggplot2)
library(ecodata)
library(lubridate)
library(dplyr)
library(stringr)
library(marmap) # bathymetry
library(RColorBrewer)
library(ggnewscale)
library(sf)
library(cowplot)
library(tidyverse)
library(ggpubr)
library(sf)
library(ggdist)
library(ggpubr)
library(wesanderson)
library(raster)
library(mgcv)
#library(glmTMB)
library(ggpmisc)

# CPUE data (no env covariates)
gt_data_model_cpue <- read.csv(here::here('data/catch_data/gt_data_model_cpue.csv'))
#names(gt_data_model_cpue) <- tolower(names(gt_data_model_cpue))
# Add in column with cpue 
# note: Paul indicated to use small mesh
gt_data_model_cpue <- gt_data_model_cpue %>% 
       rename_all(., .funs = tolower) %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM',mesh_size >= 5.6 ~ 'LG',
                                                      TRUE ~ 'NA')) %>%
  mutate(cpue_hr = sum_gt_catch/effort_dur)

# Catch data: 
sfobs <-readRDS(here::here('data/catch_data/gold_tile_sf_ob_v1_temp_price.rds'))

sfob.env <- sfobs %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM', mesh_size >= 5.6 ~ 'LG',
                              TRUE ~ 'NA'),
         cpue_hr = SUM_GT_CATCH/effort_dur) %>% 
  filter(YEAR %in% c(1998:2022) & mesh_bin == 'SM') %>%
  dplyr::select(DATE, YEAR, MONTH, YDAY,trip_id,hull_num, area, effort_dur, 
         SUM_GT_CATCH, cpue_hr, mesh_size, mesh_bin, depth, start_lat, start_lon,
         bottomT, bottomT_avg, MIN_TEMP_C, MEAN_TEMP_C, MAX_TEMP_C,
         TEMP_VARIANCE, TEMP_DEVIATION, MEAN_DPTH_M, tri, sed) %>% 
  mutate(YEAR = as.integer(YEAR)) %>% 
  rename_all(., .funs = tolower)

#write.csv(sfob.env, "C:/Users/stephanie.owen/Documents/tilefish_indicators/data/catch_data/sfob_env.csv", row.names=FALSE)


areas <- sort(unique(sfob.env$area))
catch.tally.ann <- sfob.env %>% # aggregate by year
  group_by(year) %>% 
  summarise(ttl_sum = sum(sum_gt_catch))

# Length data from observer program 
lengths <- read.csv(here::here('data/catch_data/gt_data_length_andy.csv')) 
names(lengths) <- tolower(names(lengths))

# Recruitment estimates from 2021 report
recruit <- read.csv(here::here('data/assessment_data/tilefish_rec_estimate_2021.csv'))

# Merge SF/Obs catch data with recruit estimates:
catch_recruit <- cbind(recruit %>% filter(year %in% c(1998:2020)),
                       catch.tally.ann %>%
                         filter(year %in% c(1998:2020)) %>%
                         dplyr::select(ttl_sum))

# loading in shape files for maps
wd = here::here('shapefiles')
US.areas <- st_read(here::here('shapefiles/USA.shp'), quiet = TRUE)
canada.areas <- st_read(here::here('shapefiles/Canada.shp'), quiet = TRUE)
bts_strata <- st_read(here::here('shapefiles/NES_BOTTOM_TRAWL_STRATA.shp'),
                      quiet = TRUE)
# plot(bts_strata) # to see all bottom trawl strata

gtf_strata <- bts_strata %>% 
  filter(STRATUMA %in% c('01030', '01040', '01070', '01080', '01110', '01120', 
                         '01140', '01150', '01670', '01680', '01710', '01720', 
                         '01750', '01760')) # select just the gtf strata
# plot(gtf_strata)
bathy <- marmap::getNOAA.bathy(-81,-58, 27, 46)
bathy = fortify.bathy(bathy)

```



## Tilefish data
***

### Catch data

Year-class strength is broadly defined as the number of fish spawned or hatched 
in a given year (Ricker, 1975).


Figure 1. Sum of catch (not accounting for effort), across years. Light blue 
shaded region represents the temporal range of observer records and red 
shaded region represents temporal range of study fleet records. The 'purple' 
region is where they overlap. Note that 2000-2005 for observer records had 
low sample size/number of vessels for tilefish, making the shaded region likely 
the best region to use for analysis. The vertical dashed lines represent strong
year classes for this species (Nesslage et al. 2021). Red asterisk marks year
that stock was deemed 're-built'.

```{r, c1}
# tot_catch == total (sum_catch) across hauls. so if tallying up annually, 
# use sum_catch
# Strong year-classes: 1970, 1973, 1993, 1999, 2005, 2013

ggplot(catch.tally.ann, aes(x = factor(year), y = ttl_sum, group = 1))+
   geom_rect(aes(xmin = '2007', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'red', alpha = 0.02) +
  geom_rect(aes(xmin = '2000', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'lightblue', alpha = 0.05) +
   geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.5) +
  annotate("text", label = "*",
    x = 26, y = 14000, size = 8, colour = "red" )+
  xlab('Year') + 
  ylab('Total sum tilefish catch') + 
  # facet_wrap(~month)+
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()

```


#### CPUE 


Figure 2. Catch-per-unit-effot for undirected trawl trips from the Study fleet and 
observer program. Zeros have been added using species association methodology
(via jaccard index). 

[see here for example](https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/64501a9ee391c8439cbc3c04/1682971295614/c_MRIP+Index_Drew+2022d.pdf)

```{r, c2}
gt_data_model_cpue %>% 
  filter(mesh_bin == 'SM') %>% # note: Paul indicated to use small mesh
  group_by(year, source) %>% 
  summarise(mean_cpue = mean(cpue_hr),.groups = 'drop') %>%
  ggplot(aes(x=year,y=mean_cpue)) +
  geom_line(lwd = 1) +
  facet_wrap(~source) + 
  theme_bw()

gt_data_model_cpue %>% 
  filter(mesh_bin == 'SM') %>%
  group_by(year) %>% 
  summarise(mean_cpue = mean(cpue_hr),.groups = 'drop') %>%
  ggplot(aes(x=year,y=mean_cpue)) +
  geom_line(lwd = 1) +
  labs(title = 'Study fleet + Observer combined') + 
  theme_bw()

```


#### Maps (catch) {.tabset}

Tilefish catch locations (study fleet/observer)

```{r, maps, results='asis'}

yrs = sort(unique(gt_data_model_cpue$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  gt_data_model_cpue %>% 
  filter(start_lat < 42.5 & depth_est > 50 & year == yrs) %>%
  mutate(bin = cut(year, seq(min(year), max(year) + 4, 4), right = FALSE)) %>%
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  stat_summary_2d(aes(x=start_lon, y=start_lat, z = cpue_hr),
                  binwidth=c(0.16666,0.16666)) + 
  scale_fill_viridis_c() + 
  theme(legend.position = "bottom",
        legend.key.size = unit(0.2, "cm"),
        legend.key.width = unit(1, "cm")) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'CPUE') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n#####",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```



### Recruitment estimate

Figure 3. Age-1 recruitment estimate from the 2021 tilefish assessment across 
all years

```{r, c4}

ggplot(recruit, aes(x = factor(year), y = recruit_est, group = 1))+
   geom_rect(aes(xmin = '2007', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'red', alpha = 0.02) +
  geom_rect(aes(xmin = '2000', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'lightblue', alpha = 0.05) +
   geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.5) +
  annotate("text", label = "*",
    x = 26, y = 14000, size = 8, colour = "red" )+
  xlab('Year') + 
  ylab('Total sum tilefish catch') + 
  # facet_wrap(~month)+
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()

### Thought: Should we isolate years associated w/strong year classes (or bad)
###           for correlations and analyses? 
```


Figure 4. Recruitment estimates in focus years 

+ Paul has suggested that data used to calculate recruitment 
estimates was strongest from 2000 to present. 
+ Extended to 1998 to capture both the strong year class and El nino associated 
with that year. 

```{r, c5}
ggplot(recruit %>% filter(year %in% c(1998:2022)),
       aes(x = factor(year), y = recruit_est, group = 1))+
  geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.2) +
  xlab('Year') + 
  ylab('Recruit estimates') + 
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()
```

Figure 5. Recruitment estimates and Study Fleet and Observer catch data. Black
line denotes recruitment estimate, yellow denotes sum of annual catch data
across both Study fleet and Observer programs. 

```{r, c6}
options(scipen=999)
ggplot(catch_recruit) +
  geom_line(aes(x = factor(year), y = recruit_est, group = 1),
            col = 'black', size = 1.2) +
  geom_line(aes(x = factor(year), y = ttl_sum*1000), size = 1.2, 
            color = 'goldenrod1', group = 1) +
  scale_y_continuous(sec.axis = sec_axis(~./1000, name = 'Catch (lbs)')) + 
  geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  xlab('Year') + 
  ylab('Recruit estimates') + 
   theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()
```

### Length data

#### All lengths

Figure 6. Distribution of lengths 
Figure 7. Length frequencies
Figure 8. Frequency of smaller individuals 


```{r, c7}
# Define category breaks
size_breaks <- c(0,10,20,30,40, 50, 60, 70, 80, 90, 100)
# Making a function to bin the catches
label_interval <- function(breaks) {
  paste0("(", breaks[1:length(breaks) - 1], "-", breaks[2:length(breaks)], ")")
}
labels = label_interval(size_breaks)

# length freq. table
tab = table(cut(lengths$lenanml, 
    breaks = size_breaks,
    labels = label_interval(size_breaks)))

## Plot full distribution
ggplot(lengths,
       aes(x = lenanml)) + 
  geom_bar(position = position_dodge(), 
           alpha = 0.4, fill= 'blue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  theme_facet()

# Plot length frequencies
barplot(tab, xlab = 'Length bins (mm)', main = '')

# Just the little ones
barplot(tab[1:3], xlab = 'Length bins (mm)', main = '')
```

#### Juveniles 

Young of year - year 1 and 2 size class

```{r, c8}
ggplot(lengths %>% filter(lenanml <= 26),
       aes(x = lenanml)) + 
  geom_bar(position = position_dodge(), 
           fill= 'slateblue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  theme_facet()
ggplot(lengths %>% filter(lenanml <= 26),
       aes(x = lenanml, fill = numlen)) + 
  geom_bar(position = position_dodge(), 
           alpha = 0.4, fill= 'blue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  facet_wrap(~year) + 
  theme_facet()
```


### Environmental data
***

The strong year classes for Golden Tilefish were 1993, 1998, 2005, 2013. 
Some of the underlying oceanographic processes that may be related to recruitment
may influence habitat, retention/displacement and food availablity. These are
explored below. 


#### Habitat

Tilefish occupy a very narrow band of habitat conditions. Therefore, temperature 
and salinity may be of interest. 

##### SST

```{r, c9}
# SST
sst<-read.csv(here::here('data/sst/sst_ts_gtf_strata.csv'))
```

##### Maps (SST) {.tabset}

```{r, sst maps, results='asis'}

```

##### SST analysis


```{r}
# SST by month with lag
sst.lag <- sst %>%
  mutate(mean_sst_lag2 = lag(weighted_mean_sst,2),
         mean_sst_lag3 = lag(weighted_mean_sst,3),
         mean_sst_lag6 = lag(weighted_mean_sst,6))

# filter sfob.env 
sfob <- sfob.env %>% filter(depth > 50, cpue_hr > 0) %>%
  group_by(year,month) %>% 
  summarise(mean_cpue = mean(cpue_hr)) 
         
# Join with sf/ob data
sst.sfob <- dplyr::full_join(sst.lag, sfob, by = join_by(year, month)) %>%
  dplyr::select(year, month, mean_cpue, mean_sst, weighted_mean_sst, mean_sst_lag2, mean_sst_lag3, mean_sst_lag6) %>%
  tidyr::drop_na()


## SST no lag
ggplot2::ggplot(sst.sfob,
                aes(x=mean_cpue, y=weighted_mean_sst)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Mean SST no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## SST summer
ggplot2::ggplot(sst.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=weighted_mean_sst)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Summer Mean SST') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## SST spring
ggplot2::ggplot(sst.sfob %>% filter(month %in% c(4,5,6)),
                aes(x=mean_cpue, y=weighted_mean_sst)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Spring Mean SST') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 2 month lag
ggplot2::ggplot(sst.sfob,
                aes(x=mean_cpue, y=mean_sst_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Mean SST 2 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 3 month lag
ggplot2::ggplot(sst.sfob,
                aes(x=mean_cpue, y=mean_sst_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Mean SST 3 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 6 month lag
ggplot2::ggplot(sst.sfob,
                aes(x=mean_cpue, y=mean_sst_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST') +
  labs(title = 'Mean SST 6 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```
**SST Correlation and GAM**

```{r}
# Correlation and plot
lm_sst<-lm(weighted_mean_sst ~ mean_cpue, data=sst.sfob)
summary(lm_sst)

plot(weighted_mean_sst ~ mean_cpue, data=sst.sfob, pch=1, col="dodgerblue") + abline(lm_sst)

cor.test(sst.sfob$mean_cpue, sst.sfob$weighted_mean_sst, method=c("pearson"))
ggscatter(sst.sfob, x = "mean_cpue", y = "weighted_mean_sst", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Weighted Mean SST")

# GAM
sst_gam <- gam(weighted_mean_sst ~ s(mean_cpue, k=20), data=sst.sfob, sp=0.1)
#gam.check(sst_gam)
#summary(sst_gam)

plot(sst_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(sst_gam)[1])
```


##### Bottom Temperature

Figure 1. GLORYS vs in-situ bottom temperatures from study fleet vessels. 

Figure 2. Bottom temperature (C) across years. Blue dots are in-situ data, red
dots are from GLORYS. 

```{r, c11}

ggplot2::ggplot(sfob.env, aes(x=bottomt, y=mean_temp_c)) +
  geom_point(color="blue", alpha=0.1)+
  geom_abline(intercept = 0, slope = 1) +
  xlab('Bottom Temp (SF)') +
  ylab('Bottom Temp (GLORYS)') +
  theme_bw() 

ggplot2::ggplot(sfob.env, aes(x=bottomt, y=year)) +
  geom_point(color="blue", alpha=0.1) +
  geom_point(data = sfob.env, aes(x=mean_temp_c, y=year),
             color="red", alpha=0.1) +
  xlab('Bottom Temp') +
  ylab('Year') +
  labs(color = 'Source') +
  theme_bw() 

```


##### Maps (Bottom T) {.tabset}


```{r, maps bt, results='asis'}

jet.colors <-colorRampPalette(c("blue", "#007FFF", "cyan","#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# select just years with study fleet bottom temps
sf.bt <- sfob.env %>% filter(year>2006 & depth > 50) 
yrs = sort(unique(sf.bt$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  sf.bt %>% filter(year == yrs) %>% 
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  stat_summary_2d(aes(x=start_lon, y=start_lat, z = bottomt),
                  binwidth=c(0.16666,0.16666)) + 
    scale_fill_gradientn(colors = jet.colors(20)) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'Bottom temperature (°C)') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n######",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```


##### Bottom T analysis

The following figures compare in-situ bottom temperature from the study-fleet
data set to the recruitment estimate. 

+ Note here temperatures are averaged across all depths > 50 for each month.

```{r, c13}

# Note here temperatures are averaged across all depths > 50 for each month.

# Create in-situ bottom temps by month w/lag
df.lag = sfob.env %>% filter(year > 2006 & depth > 50) %>%
  group_by(year,month) %>% 
  summarise(mean_dpth = mean(depth),
            mean_bt = mean(bottomt)) %>% 
  mutate(mean_bt_lag2 = lag(mean_bt,2),
         mean_bt_lag3 = lag(mean_bt,3), 
         mean_bt_lag6 = lag(mean_bt, 6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_dpth, 
                mean_bt, mean_bt_lag2, mean_bt_lag3, mean_bt_lag6) %>%
  tidyr::drop_na()
# See what months have data
sort(unique(df.join$month))
hist(df.join$month) # will group into spring/summer fall/winter categories
## spring/summer bottom temp no lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x=recruit_est, y=mean_bt)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp no lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x=recruit_est, y=mean_bt)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 2 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 2 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag2)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 3 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 3 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag3)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 6 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 6 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag6)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()





```



The following figures compare in-situ bottom temperature from the study-fleet
data set to the Study fleet and Observer tilefish catch data. 

+ Note here temperatures are averaged across all depths > 50 for each month.


```{r}

#filter by year, depth, cpue (includes min, max, and sd of bottomt by month)
df.lag.cpue = sfob.env %>% filter(year > 2006 & depth > 50 & cpue_hr > 0) %>%
  group_by(year,month) %>% 
  summarise(mean_dpth = mean(depth),
            mean_bt = mean(bottomt),
            min_bt = min(bottomt),
            max_bt = max(bottomt),
            sd_bt = sd(bottomt),
            mean_cpue = mean(cpue_hr)) %>% 
  mutate(mean_bt_lag2 = lag(mean_bt,2),
         mean_bt_lag3 = lag(mean_bt,3), 
         mean_bt_lag6 = lag(mean_bt, 6))

#omit NA bottomt
df.lag.sfob <- df.lag.cpue[!(is.na(df.lag.cpue$mean_bt)), ]

# See what months have data
sort(unique(df.lag.sfob$month))
hist(df.lag.sfob$month)  #winter/spring and summer/fall; summer only season with any non-zero relationship, except fall with 6 mo lag

## Winter/spring bottom temp no lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(1,2,3,4,5,6)),
                aes(x=mean_cpue, y=mean_bt)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)') +
  labs(title = 'Winter/spring no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer/fall bottom temp no lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9,10,11,12)),
                aes(x=mean_cpue, y=mean_bt)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer/fall no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer bottom temp no lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=mean_bt)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Winter/spring bottom temp 2 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(1,2,3,4,5,6)),
                aes(x = mean_cpue, y = mean_bt_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Winter/spring lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer/fall bottom temp 2 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9,10,11,12)),
                aes(x = mean_cpue, y = mean_bt_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer/fall lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer bottom temp 2 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9)),
                aes(x = mean_cpue, y = mean_bt_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Winter/spring bottom temp 3 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(1,2,3,4,5,6)),
                aes(x = mean_cpue, y = mean_bt_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Winter/spring lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer/fall bottom temp 3 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9,10,11,12)),
                aes(x = mean_cpue, y = mean_bt_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer/fall lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer bottom temp 3 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9)),
                aes(x = mean_cpue, y = mean_bt_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer/fall bottom temp 6 month lag
# No 6 month lag data for winter/spring
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9,10,11,12)),
                aes(x = mean_cpue, y = mean_bt_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer/fall lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer bottom temp 6 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(7,8,9)),
                aes(x = mean_cpue, y = mean_bt_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Summer lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Fall bottom temp 6 month lag
ggplot2::ggplot(df.lag.sfob %>% filter(month %in% c(10,11,12)),
                aes(x = mean_cpue, y = mean_bt_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Fall lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```

**Corrleation and figures for sf/ob data and in-situ bottom temps**

```{r}
lm_bt<-lm(mean_bt ~ mean_cpue, data=df.lag.sfob)
summary(lm_bt)
#plot(mean_bt ~ mean_cpue, data=df.lag.sfob, pch=1, col="dodgerblue") + abline(lm_bt)

cor.test(df.lag.sfob$mean_cpue, df.lag.sfob$mean_bt, method=c("pearson"))

ggscatter(df.lag.sfob, x = "mean_cpue", y = "mean_bt", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson", 
              label.x=20, label.y=7.5),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean bottom temp (°C)")

#ggqqplot(df.lag.sfob$mean_bt, ylab="Mean BT")
#ggqqplot(df.lag.sfob$mean_cpue, ylab="Mean CPUE")


```

**SFOB insitu bottom temp GAM**

```{r}
bt_gam <- gam(mean_bt ~ s(mean_cpue, k=50), data=df.lag.sfob, sp=0.01)
#gam.check(bt_gam)
#summary(bt_gam)

plot(bt_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(bt_gam)[1])

```


##### Salinity

Here we explore salinity from the GLORYS reanalysis model at three different 
depths

 + 55 meters (relevant to larvae, recruits)
 + 110 meters (relevant recruits, juveniles)
 + 220 meters (relevant to juveniles, adults)

```{r, c14}
# Salinity 

```

##### Maps (Salinity) {.tabset}


```{r, maps sal, results='asis'}

jet.colors <-colorRampPalette(c("blue", "#007FFF", "cyan","#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# b <- brick(here::here('data/salinity/dd_sal_55_2000_2009.tif'))
# b.00 <- b[,,,1:365]

# select just years with study fleet bottom temps
sf.bt <- sfob.env %>% filter(year>2006 & depth > 50) 
yrs = sort(unique(sf.bt$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  sf.bt %>% filter(year == yrs) %>% 
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  # stat_summary_2d(aes(x=start_lon, y=start_lat, z = bottomt),
  #                 binwidth=c(0.16666,0.16666)) + 
  #   scale_fill_gradientn(colors = jet.colors(20)) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'Salinty') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n######",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```

##### Salinity analysis

```{r}

```



#### Currents

Cross-shelf processes may influence the retention or displacement of tilefish 
during early life history stages. These are explored below.

##### Shelf water volume

Shelf water volume: A measure of the volume of water bounded inshore of the
shelf-slope front. In this analysis, shelf water is defined as all water having 
salinity <34 psu. The position of the shelf-slope front varies inter-annually
with the higher shelf water values indicating the front being pushed further 
towards the shelf break.

high shv: front pushed towards sbf
low shv: front pushed inshore (more slope water on shelf)

Hypothesis: Higher recruitment success correlated with years of higher shelf 
water volume in spring/summer. These months months may be particularly 
important as that is when spawning is occurring and the position of the sbf
may influence the position of larvae (away from spawning grounds).

Additional variables in this dataset are shelf water temperature and salinity 
which may also be indicative of habitat conditions. 


```{r, s.w.volume}

# Shelf water volume
shlfvol <- read.csv(here::here('data/shelf_water_volume/ShelfWaterVolume_BSB_update.csv'))

# wrangling date info, converting doy to date and month 
yrs <- as.vector(nrow(shlfvol))
shlfvol$Year <- as.character(shlfvol$Year)
for (i in 1:nrow(shlfvol)){
  yrs[i] <- strsplit(shlfvol$Year, ".", fixed = TRUE)[[i]][1]
}
shlfvol$year <- yrs
shlfvol <- shlfvol %>% mutate(date_= as.Date(Year.Day-1, 
                                             origin=paste0(year, "-01-01")), 
                              month= strftime(date_, "%m"), 
                              day=strftime(date_,"%d"), 
                              year = as.integer(year),
                              month = as.numeric(month))  


# Create shw vol by month w/lag
df.lag = shlfvol %>%
  group_by(year,month) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3), 
         mean_t_lag6 = lag(mean_t,6),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3), 
         mean_s_lag6 = lag(mean_s,6),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3), 
         mean_v_lag6 = lag(mean_v,6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3) %>%
  tidyr::drop_na()
# See what months have data
sort(unique(df.join$month))
hist(df.join$month) # will group into spring/summer fall/winter categories

## Shelf water volume no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()




```

With lags

**2 Months**

```{r}
## Shelf water volume 2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature  2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity 2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

```

**3 Months**


```{r}
## Shelf water volume 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


```



Annual 


```{r}

# Create shw vol by year w/lag
df.lag = shlfvol %>%
  group_by(year) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3), 
         mean_t_lag6 = lag(mean_t,6),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3), 
         mean_s_lag6 = lag(mean_s,6),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3), 
         mean_v_lag6 = lag(mean_v,6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year,recruit_est, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3,
                mean_t_lag6, mean_s_lag6, mean_v_lag6) %>%
  tidyr::drop_na()

## Shelf water vol 

ggplot2::ggplot(df.join,
                aes(x=year, y=mean_v)) + 
  geom_point(color = 'black') +
  geom_line(color = 'black') +
  xlab('Year') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  theme_bw()

ggplot2::ggplot() + 
  geom_line(data = df.join, aes(x=year, y=mean_s), color = 'red') +
  geom_line(data = df.join,aes(x=year, y=mean_t*1), color = 'blue') +
  ylim(30.0,34.0) +
  scale_y_continuous(name = 'Sh.Water Salinity', 
                      sec.axis = sec_axis(~./1, name = 'Sh.Water Temperature')) + 
  xlab('Year') +
  labs(title = 'Shelf water salinity/temperature') +
  theme_bw()

ggplot2::ggplot() + 
  geom_line(data = df.join, aes(x=year, y=mean_s), color = 'red') +
  xlab('Year') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  theme_bw()

## Shelf water vol no lag

ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


## Shelf water temperature no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


```

With lags 

```{r}

## Shelf water vol 6 yr lag

ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


## Shelf water temperature 6 yr lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water salinity 6 yr lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

```

**Shelf water volume with Study fleet/Observer data**

```{r}
# Create shw vol by month w/lag
df.lag = shlfvol %>%
  group_by(year,month) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3))

#filter sfob.env 
shlfvol.sfob = sfob.env %>% filter(depth > 50, cpue_hr > 0) %>%
  group_by(year,month) %>% 
  summarise(mean_dpth = mean(depth),
            mean_bt = mean(bottomt),
            mean_cpue = mean(cpue_hr))

# Join with sf/ob data
df.shlfvol.sfob = dplyr::full_join(shlfvol.sfob, df.lag, by = join_by(year, month)) %>%
  dplyr::select(year, month, mean_cpue, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3) %>%
  tidyr::drop_na()

#See what months have data
sort(unique(df.shlfvol.sfob$month))
hist(df.shlfvol.sfob$month) #summer and fall
```

**No lags**

```{r}
## Shelf water volume no lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume summer
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume summer') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume fall
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(10,11,12)),
                aes(x=mean_cpue, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume fall') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temp no lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_t)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water salinity no lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_s)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

ggplot2::ggplot(df.shlfvol.sfob %>% filter(year >2007 & month %in% c(7,8,9,10,11,12)),
                aes(x=mean_cpue, y=mean_v)) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  labs(title = 'SFOB x M.S.W Summer/Fall (2007:2021)') +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()

ggplot2::ggplot(df.shlfvol.sfob %>% filter(year >2007 & month %in% c(7,8,9,10,11,12)),
                aes(x=mean_cpue, y=mean_s)) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  labs(title = 'SFOB x M.S.W Summer/Fall (2007:2021)') +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()

ggplot2::ggplot(df.shlfvol.sfob %>% filter(year >2007 & month %in% c(7,8,9,10,11,12)),
                aes(x=mean_cpue, y=mean_t)) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  labs(title = 'SFOB x M.S.W Summer/Fall (2007:2021)') +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()
```

**2 month lags**

```{r}
## Shelf water volume 2 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_v_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume summer - 2 month lag
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=mean_v_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume summer - 2 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume fall - 2 month lag
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(10,11,12)),
                aes(x=mean_cpue, y=mean_v_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume fall - 2 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature 2 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_t_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water salinity 2 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_s_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```

**3 month lags**

```{r}
## Shelf water volume 3 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_v_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume summer - 3 month lag
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=mean_v_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume summer - 3 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

##Shelf water volume fall - 3 month lag
ggplot2::ggplot(df.shlfvol.sfob %>% filter(month %in% c(10,11,12)),
                aes(x=mean_cpue, y=mean_v_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume fall - 3 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature 3 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_t_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water salinity 3 month lag
ggplot2::ggplot(df.shlfvol.sfob,
                aes(x=mean_cpue, y=mean_s_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```

**Correlations between sf/ob data and shelf water volume/temp/salinity**

```{r}
lm_shlf<-lm(mean_v ~ mean_cpue, data=df.shlfvol.sfob)
summary(lm_shlf)
plot(mean_v ~ mean_cpue, data=df.shlfvol.sfob, pch=1, col="dodgerblue") + abline(lm_shlf)

cor.test(df.shlfvol.sfob$mean_cpue, df.shlfvol.sfob$mean_v, method=c("pearson"))

ggscatter(df.shlfvol.sfob, x = "mean_cpue", y = "mean_v", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean shelf water volume", title="Shelf Water Volume and CPUE Correlation")
 
# shelf water temperature
lm_shlf_t<-lm(mean_t ~ mean_cpue, data=df.shlfvol.sfob)
summary(lm_shlf_t)
plot(mean_t ~ mean_cpue, data=df.shlfvol.sfob, pch=1, col="dodgerblue") + abline(lm_shlf_t)

cor.test(df.shlfvol.sfob$mean_cpue, df.shlfvol.sfob$mean_t, method=c("pearson"))

ggscatter(df.shlfvol.sfob, x = "mean_cpue", y = "mean_t", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean shelf water temperature", title="Shelf Water Temperature and CPUE Correlation")

# shelf water salinity
lm_shlf_s<-lm(mean_s ~ mean_cpue, data=df.shlfvol.sfob)
summary(lm_shlf_s)
plot(mean_s ~ mean_cpue, data=df.shlfvol.sfob, pch=1, col="dodgerblue") + abline(lm_shlf_s)

cor.test(df.shlfvol.sfob$mean_cpue, df.shlfvol.sfob$mean_s, method=c("pearson"))

ggscatter(df.shlfvol.sfob, x = "mean_cpue", y = "mean_s", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean shelf water salinity", title="Shelf Water Salinity and CPUE Correlation")

```

**Shelf water volume GAM**

```{r}
#shelf vol GAM
shlfvol_gam <- gam(mean_v ~ s(mean_cpue, k=15), data=df.shlfvol.sfob, sp=0.01)
#gam.check(shlfvol_gam)
#summary(shlfvol_gam)

plot(shlfvol_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(shlfvol_gam)[1])

```

##### Gulf-stream index

Gulf stream index was calculated based on method presented by 
Pérez-Hernández and Joyce (2014). The gulf stream index (GSI) is a measure of 
the degrees latitude above the average Gulf Stream position based on ocean
temperature at 200m (15 C) depth between 55W to 75W.

Positive values indicate a the mean position of the GS is more Northernly, 
whereas negative values indicate a more Southernly position. 

```{r}
#gsi.m <- read.csv(here::here('C:/Users/stephanie.owen/Documents/tilefish_indicators/gulf_stream_index/mm_gsi_1954_2022_chen.csv'))

#subset only GSI, month, year and 1997>
#gsi<-gsi.m[c("GSI","year","month")] %>%
 # filter(year > 1997)
#write.csv(gsi, "C:/Users/stephanie.owen/Documents/tilefish_indicators/gulf_stream_index/gsi.csv", row.names=FALSE)

#filter sfob.env 
#gsi.sfob <- sfob.env %>% filter(depth > 50) %>%
 # group_by(year,month) %>% 
  #summarise(mean_cpue = mean(cpue_hr))
#write.csv(gsi.sfob, "C:/Users/stephanie.owen/Documents/tilefish_indicators/gulf_stream_index/gsi_sfob.csv", row.names=FALSE)

#new df wouldn't recognize month when joining the gsi and sfob data so I did it manually 
gsi <- read.csv(here::here('C:/Users/stephanie.owen/Documents/tilefish_indicators/gulf_stream_index/gsi_sfob.csv'))

gsi_nozero <- filter(gsi, mean_cpue > 0)

gsi.sfob<-mutate(gsi_nozero, pos = ifelse(GSI > 0, 'Northerly', 'Southerly'), 
         n.pos = ifelse(GSI > 0, 1, 0))

```

**GSI by Month and Season**

```{r}
# GSI all months (1998-2022)
ggplot(gsi.sfob ,aes(x=mean_cpue, y=GSI)) + 
  geom_point(color = 'black') +
  facet_wrap(~month)+
  xlab('Mean CPUE/hr') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index by month') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()

#GSI Jan-Apr
filter(gsi.sfob, month %in% c(1:4)) %>%
ggplot(.,aes(x=mean_cpue, y=GSI)) + 
 geom_point(color = 'black') +
  xlab('Mean CPUE/hr') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index (Jan:Mar)') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()

#GSI Sept-Dec
filter(gsi.sfob, month %in% c(9:12)) %>%
ggplot(.,aes(x=mean_cpue, y=GSI)) + 
  geom_point(color = 'black') +
  xlab('Mean CPUE/hr') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index (Apr:Jul)') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()

# All months
ggplot(gsi.sfob, aes(x=mean_cpue, y=GSI)) + 
  geom_point(color = 'black') +
  geom_hline(yintercept = 0, lty = 2)+
  labs(title = 'All months') +
  xlab('Mean CPUE/hr') +
  ylab('Gulf stream position anomaly') +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()

#ANOVA
pos_aov_cpue <- aov(mean_cpue ~ month * pos,
              data = gsi.sfob)

# look at effects and interactions
summary(pos_aov_cpue)
tidy_pos_aov_cpue <- broom::tidy(pos_aov_cpue)
tidy_pos_aov_cpue

ggplot(data = gsi.sfob, 
             aes(x = mean_cpue, y = GSI, fill = pos,  
                 group = year)) +
   geom_bar(color = "black", stat = "identity", 
            position = position_dodge2(preserve = "single"), width = 20) +
  theme_bw() +
  labs(title = '1998-2022',
       x = "Mean CPUE/hr", 
       y = "Gulf stream position anomaly")

```

**Correlations and GAM**

```{r}
#correlations gsi and cpue
lm_gsi<-lm(GSI ~ mean_cpue, data=gsi.sfob)
summary(lm_gsi)
plot(GSI ~ mean_cpue, data=gsi.sfob, pch=1, col="dodgerblue") + abline(lm_gsi)

cor.test(gsi.sfob$mean_cpue, gsi.sfob$GSI, method=c("pearson"))

ggscatter(gsi.sfob, x = "mean_cpue", y = "GSI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson", label.x=10, label.y=-1.0),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean Monthly GSI")

#GSI GAM
gsi_gam <- gam(GSI ~ s(mean_cpue, k=50), data=gsi.sfob, sp=0.01)
#gam.check(gsi_gam)
#summary(gsi_gam)

plot(gsi_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(gsi_gam)[1])
```

#### Food availablity 

Larval tilefish eat zooplankton, likely calanus. *Calanus finmarchicus* are a
copepod (crustacean) with a one-year life cycle and are an important food 
source for many commercially important species. Calanus spp. are lipid rich, 
herbivorous species that eat phytoplankton, diatoms in particular 
(Hobbs et al. 2020).

Diatoms are often represented as microplankton (>20 µm), but many species are of
the nanoplankton size class (2-20 µm), and a smaller few may even overlap 
with picoplanton size class (<2 µm).


##### Calanus

Calanus is not as common in MAB, need to figure out dominant zooplankton in 
MAB. 

```{r, c16}
# Calanus
calanus <- ecodata::calanus_stage %>% filter(Time %in% c(1998:2021))%>% 
    rename_all(., .funs = tolower) %>% 
   mutate(year = time)


ggplot() +
  geom_line(data = calanus %>% filter(epu == 'GB', 
                                  var == 'adt-Spring'), 
       aes(x = year , y = value, col = epu), lwd = 1) + 
  geom_line(data = calanus %>% filter(epu == 'MAB', 
                                  var == 'adt-Spring'), 
       aes(x = year , y = value, col = epu), lwd = 1) +
  labs(color = c('EPU')) +
  theme_minimal()

```


Georges Bank

```{r, c17}
# GB
c5.gb <- calanus %>% filter(epu == 'GB', var == 'c5-Spring')
adult.gb <- calanus %>% filter(epu == 'GB', var == 'adt-Spring' )

df.c5 <- dplyr::full_join(recruit, c5.gb, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()
df.adt <- dplyr::full_join(recruit, adult.gb, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()

# Regression
ggscatter(df.c5, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus c5 spring (No. per 100m^-3)",
           title = 'c5')  

ggscatter(df.adt, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus adult spring (No. per 100m^-3)",
           title = 'Adult')  
# GLM 
eqn <- as.formula(paste('recruit_est ~', paste(colnames(df.c5)[1], 
                                               collapse = " + ")))

mod0 <- glm(recruit_est ~ 1, 
            data = df.c5, 
            family = "poisson")

mod1 <- glm(eqn, 
            data = df.c5, 
            family = "poisson")

summary(mod0)
summary(mod1)
AIC(mod0, mod1) %>% dplyr::arrange(AIC)

null_prediction <- exp(predict(mod0))
mod_prediction <- exp(predict(mod1))


plot(df.c5$year, df.c5$recruit_est, type = 'l')
lines(df.c5$year, null_prediction, col = "gray")
lines(df.c5$year, mod_prediction, col = "red")

```



Mid-atlantic 

```{r, c18}
# Mid-Atlantic Bight
c5.mab <- calanus %>% filter(epu == 'MAB', var == 'c5-Spring')
adult.mab <- calanus %>% filter(epu == 'MAB', var == 'adt-Spring' )

df.c5 <- dplyr::full_join(recruit, c5.mab, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()
df.adt <- dplyr::full_join(recruit, adult.mab, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()

# Regression
ggscatter(df.c5, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus c5 spring (No. per 100m^-3)",
           title = 'c5')  

ggscatter(df.adt, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus adult spring (No. per 100m^-3)",
           title = 'Adult')  



```




##### Microplankton
```{r}
# microplankton
micro<-read.csv(here::here('data/phyto_size_class/microplankton_ts_gtf_strata.csv'))
```


##### Microplankton maps {.tabset}

```{r, micromaps, results='asis'}

```

##### Microplankton analysis

```{r}
# microplankton by month with lag
micro.lag <- micro %>%
  mutate(mean_micro_lag2 = lag(weighted_mean_micro,2),
         mean_micro_lag3 = lag(weighted_mean_micro,3),
         mean_micro_lag6 = lag(weighted_mean_micro,6))

# filter sfob.env 
sfob <- sfob.env %>% filter(depth > 50, cpue_hr > 0) %>%
  group_by(year,month) %>% 
  summarise(mean_cpue = mean(cpue_hr)) 
         
# Join with sf/ob data
micro.sfob <- dplyr::full_join(micro.lag, sfob, by = join_by(year, month)) %>%
  dplyr::select(year, month, mean_cpue, mean_micro, weighted_mean_micro, mean_micro_lag2, mean_micro_lag3, mean_micro_lag6) %>%
  tidyr::drop_na()


## Microplankton no lag
ggplot2::ggplot(micro.sfob,
                aes(x=mean_cpue, y=weighted_mean_micro)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Mean Microplankton no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Microplankton summer
ggplot2::ggplot(micro.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=weighted_mean_micro)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Summer Mean Microplankton') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Microplankton spring
ggplot2::ggplot(micro.sfob %>% filter(month %in% c(4,5,6)),
                aes(x=mean_cpue, y=weighted_mean_micro)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Spring Mean Microplankton') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 2 month lag
ggplot2::ggplot(micro.sfob,
                aes(x=mean_cpue, y=mean_micro_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Mean Microplankton 2 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 3 month lag
ggplot2::ggplot(micro.sfob,
                aes(x=mean_cpue, y=mean_micro_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Mean Microplankton 3 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 6 month lag
ggplot2::ggplot(micro.sfob,
                aes(x=mean_cpue, y=mean_micro_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean Microplankton') +
  labs(title = 'Mean Microplankton 6 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```
**Micro correlation and GAM**

```{r}
# Correlation and plot
lm_micro<-lm(weighted_mean_micro ~ mean_cpue, data=micro.sfob)
summary(lm_micro)

plot(weighted_mean_micro ~ mean_cpue, data=micro.sfob, pch=1, col="dodgerblue") + abline(lm_micro)

cor.test(micro.sfob$mean_cpue, micro.sfob$weighted_mean_micro, method=c("pearson"))
ggscatter(micro.sfob, x = "mean_cpue", y = "weighted_mean_micro", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Weighted Mean Microplankton")

# GAM
micro_gam <- gam(weighted_mean_micro ~ s(mean_cpue, k=20), data=micro.sfob, sp=0.01)
#gam.check(micro_gam)
#summary(micro_gam)

plot(micro_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(micro_gam)[1])
```




##### Chlorophyll-A
```{r}

# CHL-A
chl<-read.csv(here::here('data/chl/chl_ts_gtf_strata.csv'))
```

##### Maps (Chlorophyll-A) {.tabset}
```{r, chlmaps, results='asis'}

```

##### Chlorophyll-A analysis

```{r, c21}
# chl by month with lag
chl.lag <- chl %>%
  mutate(mean_chl_lag2 = lag(weighted_mean_chl,2),
         mean_chl_lag3 = lag(weighted_mean_chl,3),
         mean_chl_lag6 = lag(weighted_mean_chl,6))

# filter sfob.env 
sfob <- sfob.env %>% filter(depth > 50, cpue_hr > 0) %>%
  group_by(year,month) %>% 
  summarise(mean_cpue = mean(cpue_hr)) 
         
# Join with sf/ob data
chl.sfob <- dplyr::full_join(chl.lag, sfob, by = join_by(year, month)) %>%
  dplyr::select(year, month, mean_cpue, mean_chl, weighted_mean_chl, mean_chl_lag2, mean_chl_lag3, mean_chl_lag6) %>%
  tidyr::drop_na()


## CHL-a no lag
ggplot2::ggplot(chl.sfob,
                aes(x=mean_cpue, y=weighted_mean_chl)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Mean CHL-a no lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## CHL summer
ggplot2::ggplot(chl.sfob %>% filter(month %in% c(7,8,9)),
                aes(x=mean_cpue, y=weighted_mean_chl)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Summer Mean CHL-a') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## CHL fall
ggplot2::ggplot(chl.sfob %>% filter(month %in% c(10,11,12)),
                aes(x=mean_cpue, y=weighted_mean_chl)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Fall Mean CHL-a') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 2 month lag
ggplot2::ggplot(chl.sfob,
                aes(x=mean_cpue, y=mean_chl_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Mean CHL-a 2 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 3 month lag
ggplot2::ggplot(chl.sfob,
                aes(x=mean_cpue, y=mean_chl_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Mean CHL-a 3 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## 6 month lag
ggplot2::ggplot(chl.sfob,
                aes(x=mean_cpue, y=mean_chl_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean CHL-a') +
  labs(title = 'Mean CHL-a 6 month lag') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
```

**CHL-a and CPUE correlation**

```{r}
lm_chl<-lm(weighted_mean_chl ~ mean_cpue, data=chl.sfob)
summary(lm_chl)

plot(weighted_mean_chl ~ mean_cpue, data=chl.sfob, pch=1, col="dodgerblue") + abline(lm_chl)

cor.test(chl.sfob$mean_cpue, chl.sfob$weighted_mean_chl, method=c("pearson"))
ggscatter(chl.sfob, x = "mean_cpue", y = "weighted_mean_chl", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean Monthly CHL-a")
```
**CHL-a GAM**

```{r}
chl_gam <- gam(weighted_mean_chl ~ s(mean_cpue, k=20), data=chl.sfob, sp=0.01)
#gam.check(chl_gam)
#summary(chl_gam)

plot(chl_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(chl_gam)[1])
```



##### SST Fronts

```{r}
# SST fronts
fprob<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_3x3.csv')) #fprob across all individual strata
fprob.ind<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_indv_substrata_3x3.csv')) #mean for each of 14 substrata
frob.ns<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_n_v_s_substrata_3x3.csv')) #mean for substrata N and S of Hudson canyon

```

##### Maps (SST fronts) {.tabset}
```{r, sst front maps, results='asis'}


```

##### SST Fronts analysis

**FPROB across all individual strata**
```{r}
#sfob by season
sfob_season<-read.csv(here::here('data/sst_fronts/sfob_byseason.csv'))
         
#join with fprob
fprob.sfob <- dplyr::full_join(fprob, sfob_season, by = join_by(year, season)) %>%
  dplyr::select(year, season, mean_cpue, mean_fprob, weighted_mean_fprob) %>%
  tidyr::drop_na()

## All months
ggplot2::ggplot(fprob.sfob,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Mean SST frontal probability') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Winter
ggplot2::ggplot(fprob.sfob %>% filter(season == "winter") ,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Winter Mean SST frontal probability') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Spring
ggplot2::ggplot(fprob.sfob %>% filter(season == "spring"),
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Spring Mean SST frontal probability') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Summer
ggplot2::ggplot(fprob.sfob %>% filter(season == "summer"),
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Summer Mean SST frontal probability') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Fall
ggplot2::ggplot(fprob.sfob %>% filter(season == "fall"),
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Fall Mean SST frontal probability') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Correlation and GAM
lm_fprob<-lm(weighted_mean_fprob ~ mean_cpue, data=fprob.sfob)
summary(lm_fprob)

plot(weighted_mean_fprob ~ mean_cpue, data=fprob.sfob, pch=1, col="dodgerblue") + abline(lm_fprob)

cor.test(fprob.sfob$mean_cpue, fprob.sfob$weighted_mean_fprob, method=c("pearson"))
ggscatter(fprob.sfob, x = "mean_cpue", y = "weighted_mean_fprob", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean Seasonal SST Frontal probability")

fprob_gam <- gam(weighted_mean_fprob ~ s(mean_cpue, k=30), data=fprob.sfob, sp=0.001)
gam.check(fprob_gam)
summary(fprob_gam)

plot(fprob_gam, residuals = TRUE, pch = 1, cex=0.5, shade=TRUE, shade.col="lightblue", seWithMean = TRUE, shift = coef(fprob_gam)[1])

```

**FPROB of each substrata**

```{r}
#join with fprob
ind.sfob <- dplyr::full_join(fprob.ind, sfob_season, by = join_by(year, season)) %>%
  dplyr::select(year, season, id, mean_cpue, mean_fprob, weighted_mean_fprob) %>%
  tidyr::drop_na()

# fprob by substrata
ggplot2::ggplot(ind.sfob,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Mean SST frontal probability') +
  stat_poly_eq(aes(label = paste(..rr.label.., sep = "~~~")), 
               label.x.npc = 30, label.y.npc = 0.15,
               parse = TRUE, size = 3)  +
  theme_bw() +
  facet_wrap(~id) +
  theme_facet()

## Winter
ggplot2::ggplot(ind.sfob %>% filter(season == "winter") ,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Winter Mean SST frontal probability') +
  stat_poly_eq(aes(label = paste(..rr.label.., sep = "~~~")), 
               label.x.npc = 30, label.y.npc = 0.15,
               parse = TRUE, size = 3)  +
  theme_bw() +
  facet_wrap(~id) +
  theme_facet()

## Spring
ggplot2::ggplot(ind.sfob %>% filter(season == "spring") ,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Spring Mean SST frontal probability') +
  stat_poly_eq(aes(label = paste(..rr.label.., sep = "~~~")), 
               label.x.npc = 30, label.y.npc = 0.15,
               parse = TRUE, size = 3)  +
  theme_bw() +
  facet_wrap(~id) +
  theme_facet()

## Summer
ggplot2::ggplot(ind.sfob %>% filter(season == "summer") ,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Summer Mean SST frontal probability') +
  stat_poly_eq(aes(label = paste(..rr.label.., sep = "~~~")), 
               label.x.npc = 30, label.y.npc = 0.15,
               parse = TRUE, size = 3)  +
  theme_bw() +
  facet_wrap(~id) +
  theme_facet()

## Fall
ggplot2::ggplot(ind.sfob %>% filter(season == "fall") ,
                aes(x=mean_cpue, y=weighted_mean_fprob)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Mean CPUE/hr') +
  ylab('Weighted mean SST frontal probability') +
  labs(title = 'Fall Mean SST frontal probability') +
  stat_poly_eq(aes(label = paste(..rr.label.., sep = "~~~")), 
               label.x.npc = 30, label.y.npc = 0.15,
               parse = TRUE, size = 3)  +
  theme_bw() +
  facet_wrap(~id) +
  theme_facet()

## Scatter plot with correlations
ggscatter(ind.sfob, x = "mean_cpue", y = "weighted_mean_fprob", 
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.coeff.args = list(method="pearson", 
              label.x=20, label.y=0.6, size =3),
          add.params=list(color="dodgerblue", fill="lightgray"),
          xlab = "Mean CPUE/hr", ylab = "Mean SST frontal probability") +
  facet_wrap(~id) +
  theme_facet()
```





References: 


Hobbs, L., Banas, N. S., Cottier, F. R., Berge, J., & Daase, M. (2020). Eat or 
sleep: availability of winter prey explains mid-winter and spring activity in
an Arctic Calanus population. Frontiers in Marine Science, 7, 541564.













