---
title: "Environmental indicator analysis-Tilefish"
output:  
  html_document:
    code_fold: hide
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, wrangling}

# Loading packages and data
library(ggplot2)
library(ecodata)
library(lubridate)
library(dplyr)
library(stringr)
library(marmap) # bathymetry
library(RColorBrewer)
library(ggnewscale)
library(sf)
library(cowplot)
library(tidyverse)
library(ggpubr)
library(sf)
library(ggdist)
library(ggpubr)
library(wesanderson)
library(raster)
#library(glmTMB)

# CPUE data (no env covariates)
gt_data_model_cpue <- read.csv(here::here('data/catch_data/gt_data_model_cpue.csv'))
#names(gt_data_model_cpue) <- tolower(names(gt_data_model_cpue))
# Add in column with cpue 
# note: Paul indicated to use small mesh
gt_data_model_cpue <- gt_data_model_cpue %>% 
       rename_all(., .funs = tolower) %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM',mesh_size >= 5.6 ~ 'LG',
                                                      TRUE ~ 'NA')) %>%
  mutate(cpue_hr = sum_gt_catch/effort_dur)

# Catch data: 
sfobs <-readRDS(here::here('data/catch_data/gold_tile_sf_ob_v1_temp_price.rds'))

sfob.env <- sfobs %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM', mesh_size >= 5.6 ~ 'LG',
                              TRUE ~ 'NA'),
         cpue_hr = SUM_GT_CATCH/effort_dur) %>% 
  filter(YEAR %in% c(1998:2022) & mesh_bin == 'SM') %>%
  dplyr::select(DATE, YEAR, MONTH, YDAY,trip_id,hull_num, area, effort_dur, 
         SUM_GT_CATCH, cpue_hr, mesh_size, mesh_bin, depth, start_lat, start_lon,
         bottomT, bottomT_avg, MIN_TEMP_C, MEAN_TEMP_C, MAX_TEMP_C,
         TEMP_VARIANCE, TEMP_DEVIATION, MEAN_DPTH_M, tri, sed) %>% 
  mutate(YEAR = as.integer(YEAR)) %>% 
  rename_all(., .funs = tolower)


areas <- sort(unique(sfob.env$area))
catch.tally.ann <- sfob.env %>% # aggregate by year
  group_by(year) %>% 
  summarise(ttl_sum = sum(sum_gt_catch))

# Length data from observer program 
lengths <- read.csv(here::here('data/catch_data/gt_data_length_andy.csv')) 
names(lengths) <- tolower(names(lengths))

# Recruitment estimates from 2021 report
recruit <- read.csv(here::here('data/assessment_data/tilefish_rec_estimate_2021.csv'))

# Merge SF/Obs catch data with recruit estimates:
catch_recruit <- cbind(recruit %>% filter(year %in% c(1998:2020)),
                       catch.tally.ann %>%
                         filter(year %in% c(1998:2020)) %>%
                         dplyr::select(ttl_sum))

# loading in shape files for maps
wd = here::here('shapefiles')
US.areas <- st_read(here::here('shapefiles/USA.shp'), quiet = TRUE)
canada.areas <- st_read(here::here('shapefiles/Canada.shp'), quiet = TRUE)
bts_strata <- st_read(here::here('shapefiles/NES_BOTTOM_TRAWL_STRATA.shp'),
                      quiet = TRUE)
# plot(bts_strata) # to see all bottom trawl strata

gtf_strata <- bts_strata %>% 
  filter(STRATUMA %in% c('01030', '01040', '01070', '01080', '01110', '01120', 
                         '01140', '01150', '01670', '01680', '01710', '01720', 
                         '01750', '01760')) # select just the gtf strata
# plot(gtf_strata)
bathy <- marmap::getNOAA.bathy(-81,-58, 27, 46)
bathy = fortify.bathy(bathy)

```



## Tilefish data
***

### Catch data

Year-class strength is broadly defined as the number of fish spawned or hatched 
in a given year (Ricker, 1975).


Figure 1. Sum of catch (not accounting for effort), across years. Light blue 
shaded region represents the temporal range of observer records and red 
shaded region represents temporal range of study fleet records. The 'purple' 
region is where they overlap. Note that 2000-2005 for observer records had 
low sample size/number of vessels for tilefish, making the shaded region likely 
the best region to use for analysis. The vertical dashed lines represent strong
year classes for this species (Nesslage et al. 2021). Red asterisk marks year
that stock was deemed 're-built'.

```{r, c1}
# tot_catch == total (sum_catch) across hauls. so if tallying up annually, 
# use sum_catch
# Strong year-classes: 1970, 1973, 1993, 1999, 2005, 2013

ggplot(catch.tally.ann, aes(x = factor(year), y = ttl_sum, group = 1))+
   geom_rect(aes(xmin = '2007', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'red', alpha = 0.02) +
  geom_rect(aes(xmin = '2000', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'lightblue', alpha = 0.05) +
   geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.5) +
  annotate("text", label = "*",
    x = 26, y = 14000, size = 8, colour = "red" )+
  xlab('Year') + 
  ylab('Total sum tilefish catch') + 
  # facet_wrap(~month)+
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()

```


#### CPUE 


Figure 2. Catch-per-unit-effot for undirected trawl trips from the Study fleet and 
observer program. Zeros have been added using species association methodology
(via jaccard index). 

[see here for example](https://static1.squarespace.com/static/511cdc7fe4b00307a2628ac6/t/64501a9ee391c8439cbc3c04/1682971295614/c_MRIP+Index_Drew+2022d.pdf)

```{r, c2}
gt_data_model_cpue %>% 
  filter(mesh_bin == 'SM') %>% # note: Paul indicated to use small mesh
  group_by(year, source) %>% 
  summarise(mean_cpue = mean(cpue_hr),.groups = 'drop') %>%
  ggplot(aes(x=year,y=mean_cpue)) +
  geom_line(lwd = 1) +
  facet_wrap(~source) + 
  theme_bw()

gt_data_model_cpue %>% 
  filter(mesh_bin == 'SM') %>%
  group_by(year) %>% 
  summarise(mean_cpue = mean(cpue_hr),.groups = 'drop') %>%
  ggplot(aes(x=year,y=mean_cpue)) +
  geom_line(lwd = 1) +
  labs(title = 'Study fleet + Observer combined') + 
  theme_bw()

```


#### Maps (catch) {.tabset}

Tilefish catch locations (study fleet/observer)

```{r, maps, results='asis'}

yrs = sort(unique(gt_data_model_cpue$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  gt_data_model_cpue %>% 
  filter(start_lat < 42.5 & depth_est > 50 & year == yrs) %>%
  mutate(bin = cut(year, seq(min(year), max(year) + 4, 4), right = FALSE)) %>%
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  stat_summary_2d(aes(x=start_lon, y=start_lat, z = cpue_hr),
                  binwidth=c(0.16666,0.16666)) + 
  scale_fill_viridis_c() + 
  theme(legend.position = "bottom",
        legend.key.size = unit(0.2, "cm"),
        legend.key.width = unit(1, "cm")) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'CPUE') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n#####",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```



### Recruitment estimate

Figure 3. Age-1 recruitment estimate from the 2021 tilefish assessment across 
all years

```{r, c4}

ggplot(recruit, aes(x = factor(year), y = recruit_est, group = 1))+
   geom_rect(aes(xmin = '2007', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'red', alpha = 0.02) +
  geom_rect(aes(xmin = '2000', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'lightblue', alpha = 0.05) +
   geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.5) +
  annotate("text", label = "*",
    x = 26, y = 14000, size = 8, colour = "red" )+
  xlab('Year') + 
  ylab('Total sum tilefish catch') + 
  # facet_wrap(~month)+
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()

### Thought: Should we isolate years associated w/strong year classes (or bad)
###           for correlations and analyses? 
```


Figure 4. Recruitment estimates in focus years 

+ Paul has suggested that data used to calculate recruitment 
estimates was strongest from 2000 to present. 
+ Extended to 1998 to capture both the strong year class and El nino associated 
with that year. 

```{r, c5}
ggplot(recruit %>% filter(year %in% c(1998:2022)),
       aes(x = factor(year), y = recruit_est, group = 1))+
  geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.2) +
  xlab('Year') + 
  ylab('Recruit estimates') + 
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()
```

Figure 5. Recruitment estimates and Study Fleet and Observer catch data. Black
line denotes recruitment estimate, yellow denotes sum of annual catch data
across both Study fleet and Observer programs. 

```{r, c6}
options(scipen=999)
ggplot(catch_recruit) +
  geom_line(aes(x = factor(year), y = recruit_est, group = 1),
            col = 'black', size = 1.2) +
  geom_line(aes(x = factor(year), y = ttl_sum*1000), size = 1.2, 
            color = 'goldenrod1', group = 1) +
  scale_y_continuous(sec.axis = sec_axis(~./1000, name = 'Catch (lbs)')) + 
  geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  xlab('Year') + 
  ylab('Recruit estimates') + 
   theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()
```

### Length data

#### All lengths

Figure 6. Distribution of lengths 
Figure 7. Length frequencies
Figure 8. Frequency of smaller individuals 


```{r, c7}
# Define category breaks
size_breaks <- c(0,10,20,30,40, 50, 60, 70, 80, 90, 100)
# Making a function to bin the catches
label_interval <- function(breaks) {
  paste0("(", breaks[1:length(breaks) - 1], "-", breaks[2:length(breaks)], ")")
}
labels = label_interval(size_breaks)

# length freq. table
tab = table(cut(lengths$lenanml, 
    breaks = size_breaks,
    labels = label_interval(size_breaks)))

## Plot full distribution
ggplot(lengths,
       aes(x = lenanml)) + 
  geom_bar(position = position_dodge(), 
           alpha = 0.4, fill= 'blue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  theme_facet()

# Plot length frequencies
barplot(tab, xlab = 'Length bins (mm)', main = '')

# Just the little ones
barplot(tab[1:3], xlab = 'Length bins (mm)', main = '')
```

#### Juveniles 

Young of year - year 1 and 2 size class

```{r, c8}
ggplot(lengths %>% filter(lenanml <= 26),
       aes(x = lenanml)) + 
  geom_bar(position = position_dodge(), 
           fill= 'slateblue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  theme_facet()
ggplot(lengths %>% filter(lenanml <= 26),
       aes(x = lenanml, fill = numlen)) + 
  geom_bar(position = position_dodge(), 
           alpha = 0.4, fill= 'blue', color="black") + 
  xlab('Tilefish length (cm)') +
  theme_bw() +
  facet_wrap(~year) + 
  theme_facet()
```


### Environmental data
***

The strong year classes for Golden Tilefish were 1993, 1998, 2005, 2013. 
Some of the underlying oceanographic processes that may be related to recruitment
may influence habitat, retention/displacement and food availablity. These are
explored below. 


#### Habitat

Tilefish occupy a very narrow band of habitat conditions. Therefore, temperature 
and salinity may be of interest. 

##### SST

```{r, c9}
# SST

```

##### Maps (SST) {.tabset}

```{r, sst maps, results='asis'}

```

##### SST analysis


```{r}

```


##### Bottom Temperature

Figure 1. GLORYS vs in-situ bottom temperatures from study fleet vessels. 

Figure 2. Bottom temperature (C) across years. Blue dots are in-situ data, red
dots are from GLORYS. 

```{r, c11}

ggplot2::ggplot(sfob.env, aes(x=bottomt, y=mean_temp_c)) +
  geom_point(color="blue", alpha=0.1)+
  geom_abline(intercept = 0, slope = 1) +
  xlab('Bottom Temp (SF)') +
  ylab('Bottom Temp (GLORYS)') +
  theme_bw() 

ggplot2::ggplot(sfob.env, aes(x=bottomt, y=year)) +
  geom_point(color="blue", alpha=0.1) +
  geom_point(data = sfob.env, aes(x=mean_temp_c, y=year),
             color="red", alpha=0.1) +
  xlab('Bottom Temp') +
  ylab('Year') +
  labs(color = 'Source') +
  theme_bw() 

```


##### Maps (Bottom T) {.tabset}


```{r, maps bt, results='asis'}

jet.colors <-colorRampPalette(c("blue", "#007FFF", "cyan","#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# select just years with study fleet bottom temps
sf.bt <- sfob.env %>% filter(year>2006 & depth > 50) 
yrs = sort(unique(sf.bt$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  sf.bt %>% filter(year == yrs) %>% 
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  stat_summary_2d(aes(x=start_lon, y=start_lat, z = bottomt),
                  binwidth=c(0.16666,0.16666)) + 
    scale_fill_gradientn(colors = jet.colors(20)) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'Bottom temperature (°C)') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n######",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```


##### Bottom T analysis

The following figures compare in-situ bottom temperature from the study-fleet
data set to the recruitment estimate. 

+ Note here temperatures are averaged across all depths > 50 for each month.

```{r, c13}

# Note here temperatures are averaged across all depths > 50 for each month.

# Create in-situ bottom temps by month w/lag
df.lag = sfob.env %>% filter(year > 2006 & depth > 50) %>%
  group_by(year,month) %>% 
  summarise(mean_dpth = mean(depth),
            mean_bt = mean(bottomt)) %>% 
  mutate(mean_bt_lag2 = lag(mean_bt,2),
         mean_bt_lag3 = lag(mean_bt,3), 
         mean_bt_lag6 = lag(mean_bt, 6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_dpth, 
                mean_bt, mean_bt_lag2, mean_bt_lag3, mean_bt_lag6) %>%
  tidyr::drop_na()
# See what months have data
sort(unique(df.join$month))
hist(df.join$month) # will group into spring/summer fall/winter categories
## spring/summer bottom temp no lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x=recruit_est, y=mean_bt)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp no lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x=recruit_est, y=mean_bt)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 2 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 2 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag2)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 3 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 3 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag3)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 6 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(4,5,6,7,8)),
                aes(x = recruit_est, y = mean_bt_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## spring/summer bottom temp 6 month lag
ggplot2::ggplot(df.join %>% filter(month %in% c(9,10,11,12)),
                aes(x = recruit_est, y = mean_bt_lag6)) + 
  geom_point(color= 'black')+
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean bottom temp (°C)')+
  labs(title = 'Lag 6 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()





```



The following figures compare in-situ bottom temperature from the study-fleet
data set to the Study fleet and Observer tilefish catch data. 

+ Note here temperatures are averaged across all depths > 50 for each month.

**Stephanie I hope the code below is a good jumping off point!** 

```{r}
# Stephanie here I subset the years that are relevant to the Study fleet 
# data set. Note that Paul suggested looking at catch that is from depths 
# greater than 50 m and only from small mesh gear. The object sfob.env is
# already filtered for just small gear (see line 57 above).


# You can run this line just to check. 
unique(sfob.env$mesh_bin)

# the lines that create the df.lag object selects  depths > 50, just note 
# that there are many depths here, so this is taking the average temperature
# across all depths. We may want to think a bit more about how we want to 
# subset for averaging (or not?)

# here is a way to see what depths are reported 
sort(unique(sfob.env$depth))

# another note - since I group by year AND month - the lags are month specific. 
# If you wanted to lag by year only, you would have to get rid of month below.
df.lag = sfob.env %>% filter(year > 2006 & depth > 50) %>%
  group_by(year,month) %>% 
  summarise(mean_dpth = mean(depth),
            mean_bt = mean(bottomt)) %>% 
  mutate(mean_bt_lag2 = lag(mean_bt,2),
         mean_bt_lag3 = lag(mean_bt,3), 
         mean_bt_lag6 = lag(mean_bt, 6))



```



##### Salinity

Here we explore salinity from the GLORYS reanalysis model at three different 
depths

 + 55 meters (relevant to larvae, recruits)
 + 110 meters (relevant recruits, juveniles)
 + 220 meters (relevant to juveniles, adults)

```{r, c14}
# Salinity 

```

##### Maps (Salinity) {.tabset}


```{r, maps sal, results='asis'}

jet.colors <-colorRampPalette(c("blue", "#007FFF", "cyan","#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

# b <- brick(here::here('data/salinity/dd_sal_55_2000_2009.tif'))
# b.00 <- b[,,,1:365]

# select just years with study fleet bottom temps
sf.bt <- sfob.env %>% filter(year>2006 & depth > 50) 
yrs = sort(unique(sf.bt$year))    

#for(i in 1:length(yrs)){
yrmap <- function(yrs){
  sf.bt %>% filter(year == yrs) %>% 
  ggplot() + 
  geom_sf(data = US.areas %>% st_as_sf(),color = 'gray20', fill = '#cbdbcd') +
  geom_contour(data = bathy,
               aes(x=x,y=y,z=-1*z),
               breaks=c(50,100,150,200, Inf),
               size=c(0.3),
               col = 'darkgrey') +
  # stat_summary_2d(aes(x=start_lon, y=start_lat, z = bottomt),
  #                 binwidth=c(0.16666,0.16666)) + 
  #   scale_fill_gradientn(colors = jet.colors(20)) +
  coord_sf(xlim = c(-75,-65.5), ylim = c(36,44), datum = sf::st_crs(4326))  +
  labs(x = '', y = '', fill = 'Salinty') +
  theme_bw() 
}


for(i in 1:length(yrs)){
 cat("\n######",  as.character(yrs[i]),"\n")
    print(yrmap(yrs[i])) 
    cat("\n")   
}


```

##### Salinity analysis

```{r}

```



#### Currents

Cross-shelf processes may influence the retention or displacement of tilefish 
during early life history stages. These are explored below.

##### Shelf water volume

Shelf water volume: A measure of the volume of water bounded inshore of the
shelf-slope front. In this analysis, shelf water is defined as all water having 
salinity <34 psu. The position of the shelf-slope front varies inter-annually
with the higher shelf water values indicating the front being pushed further 
towards the shelf break.

high shv: front pushed towards sbf
low shv: front pushed inshore (more slope water on shelf)

Hypothesis: Higher recruitment success correlated with years of higher shelf 
water volume in spring/summer. These months months may be particularly 
important as that is when spawning is occurring and the position of the sbf
may influence the position of larvae (away from spawning grounds).

Additional variables in this dataset are shelf water temperature and salinity 
which may also be indicative of habitat conditions. 


```{r, s.w.volume}

# Shelf water volume
shlfvol <- read.csv(here::here('data/shelf_water_volume/ShelfWaterVolume_BSB_update.csv'))

# wrangling date info, converting doy to date and month 
yrs <- as.vector(nrow(shlfvol))
shlfvol$Year <- as.character(shlfvol$Year)
for (i in 1:nrow(shlfvol)){
  yrs[i] <- strsplit(shlfvol$Year, ".", fixed = TRUE)[[i]][1]
}
shlfvol$year <- yrs
shlfvol <- shlfvol %>% mutate(date_= as.Date(Year.Day-1, 
                                             origin=paste0(year, "-01-01")), 
                              month= strftime(date_, "%m"), 
                              day=strftime(date_,"%d"), 
                              year = as.integer(year),
                              month = as.numeric(month))  


# Create shw vol by month w/lag
df.lag = shlfvol %>%
  group_by(year,month) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3), 
         mean_t_lag6 = lag(mean_t,6),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3), 
         mean_s_lag6 = lag(mean_s,6),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3), 
         mean_v_lag6 = lag(mean_v,6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3) %>%
  tidyr::drop_na()
# See what months have data
sort(unique(df.join$month))
hist(df.join$month) # will group into spring/summer fall/winter categories
## Shelf water volume no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


ggplot2::ggplot(df.join %>% filter(year >1997 & month %in% c(7,8,9)),
                aes(x=recruit_est, y=mean_s)) + 
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  labs(title = 'Recruitment est x M.S.W July,Aug,Sept (1998:2020)') +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()

```

With lags

**2 Months**

```{r}
## Shelf water volume 2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature  2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity 2 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag2)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 2 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

```

**3 Months**


```{r}
## Shelf water volume 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water temperature 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity 3 month lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag3)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 3 months') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


```



Annual 


```{r}

# Create shw vol by year w/lag
df.lag = shlfvol %>%
  group_by(year) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3), 
         mean_t_lag6 = lag(mean_t,6),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3), 
         mean_s_lag6 = lag(mean_s,6),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3), 
         mean_v_lag6 = lag(mean_v,6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year,recruit_est, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3,
                mean_t_lag6, mean_s_lag6, mean_v_lag6) %>%
  tidyr::drop_na()

## Shelf water vol 

ggplot2::ggplot(df.join,
                aes(x=year, y=mean_v)) + 
  geom_point(color = 'black') +
  geom_line(color = 'black') +
  xlab('Year') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  theme_bw()

ggplot2::ggplot() + 
  geom_line(data = df.join, aes(x=year, y=mean_s), color = 'red') +
  geom_line(data = df.join,aes(x=year, y=mean_t*1), color = 'blue') +
  ylim(30.0,34.0) +
  scale_y_continuous(name = 'Sh.Water Salinity', 
                      sec.axis = sec_axis(~./1, name = 'Sh.Water Temperature')) + 
  xlab('Year') +
  labs(title = 'Shelf water salinity/temperature') +
  theme_bw()

ggplot2::ggplot() + 
  geom_line(data = df.join, aes(x=year, y=mean_s), color = 'red') +
  xlab('Year') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  theme_bw()

## Shelf water vol no lag

ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


## Shelf water temperature no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()
## Shelf water salinity no lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


```

With lags 

```{r}

## Shelf water vol 6 yr lag

ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_v_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water volume') +
  labs(title = 'Shelf water volume - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()


## Shelf water temperature 6 yr lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_t_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water temperature') +
  labs(title = 'Shelf water temperature - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

## Shelf water salinity 6 yr lag
ggplot2::ggplot(df.join,
                aes(x=recruit_est, y=mean_s_lag6)) + 
  geom_point(color = 'black') +
  stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth") +
  xlab('Recruitment estimate') +
  ylab('Mean shelf water salinity') +
  labs(title = 'Shelf water salinity - lag 6 years') +
  stat_cor(aes(label=..rr.label..)) +
  theme_bw()

```


##### Gulf-stream index

Gulf stream index was calculated based on method presented by 
Pérez-Hernández and Joyce (2014). The gulf stream index (GSI) is a measure of 
the degrees latitude above the average Gulf Stream position based on ocean
temperature at 200m (15 C) depth between 55W to 75W.

Positive values indicate a the mean position of the GS is more Northernly, 
whereas negative values indicate a more Southernly position. 

```{r}
# positive are more Northerly, negative are more southernly 

## --- Note below is the original data source, which was modified to include 
## --- month and then saved as a new file (mm_gsi_1954_2022_chen.csv).
# gsi.a <- ecodata::gsi
# gsi.m <- read.csv(here::here('data/gulf_stream_index/Chen_EN4_T200_GSI_1954_2022_monthly - Zhuomin Chen.xlsx - Sheet1.csv'))

# # weird inefficient solution I came up with to get month values (but it works)
# gsi.m$year <- as.numeric(gsub("(^\\d{4}).*", "\\1", gsi.m$Month))
# gsi.m$m.1 <- round(str_extract(gsi.m$Month, '\\d+([.,]\\d+)?') %>% as.numeric()- gsi.m$year,2)
# gsi.m$month <- round(as.numeric(str_extract(gsi.m$m.1, '\\d\\d')))
# is.na(gsi.m$month) <- 10
# gsi.m$month <- replace(gsi.m$month, is.na(gsi.m$month), 10)
# gsi.m <- gsi.m[,-4]
# # saving so don't have to do that everytime:
# write.csv(gsi.m, 'mm_gsi_1954_2022_chen.csv') #


gsi.m <- read.csv(here::here('data/gulf_stream_index/mm_gsi_1954_2022_chen.csv'))

# df <- dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
#   dplyr::select(year, month, recruit_est, GSI) %>%
#   filter(month %in% c(3:9)) %>% 
#   tidyr::pivot_wider(names_from = c(month),
#                      values_from = c(GSI)) %>% 
#   rlang::set_names(c('year','recruit_est','Mar', 'Apr', 'May',                                                                   'Jun', 'Jul', 'Aug', 'Sep')) %>% 
#   tidyr::drop_na()


## --- This joins recruit estimate data to GSI data and 
## --- calculates some summary stats(mean, sd, min, max)
df <- dplyr::full_join(recruit, gsi.m %>%
                   group_by(year) %>% 
                   filter(month %in% c(3:8)) %>% 
                   summarise(m.gsi = mean(GSI),
                             sd.gsi = sd(GSI),
                             max.gsi = max(GSI), 
                             min.gsi = min(GSI)), 
                 by = join_by(year)) %>% 
  mutate(pos = ifelse(m.gsi > 0, 'Northerly', 'Southerly'), 
         n.pos = ifelse(m.gsi > 0, 1, 0))
df <- df[-c(51:69),] # this removes rows w/years that don't match 
# (could/should do this before hand in more standardized way - will do later)

# Plot the time series (mean GSI for the months of interest)
# -- Note: here I am looking just at March through August assuming these
# months are the most important to recently spawned individuals 
ggplot(data = df, 
       aes(x = year, y = m.gsi)) +
  geom_line(lwd = 1) +
  geom_hline(yintercept = 0, lty = 2) +
  labs(title = 'Mean GS position anomaly March:August',
       x = 'Year', 
       y = "Gulf stream position anomaly\n") +
theme_bw() 

# dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
#   dplyr::select(year, month, recruit_est, GSI) %>%
#   #filter(month %in% c(3:9)) %>%  
#   filter(year>1997) %>% 
#   tidyr::drop_na() %>% 
#   ggplot2::ggplot(.,aes(x=recruit_est, y=GSI)) + 
#   geom_point(color = 'black') +
#   xlab('Recruitment estimate') +
#   ylab('Gulf stream position anomaly') +
#   labs(title = 'Gulf stream index') +
#   geom_hline(yintercept = 0, lty = 2) +
#   theme_bw()


# Looking across all months 
dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, GSI) %>%
  #filter(month %in% c(3:9)) %>%  
  filter(year>1997) %>% 
  tidyr::drop_na() %>% 
ggplot2::ggplot(.,aes(x=recruit_est, y=GSI)) + 
  geom_point(color = 'black') +
  facet_wrap(~month)+
  xlab('Recruitment estimate') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index by month') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()

# Looking across April through July
dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, GSI) %>%
  #filter(month %in% c(3:9)) %>%  
  filter(year>1997 & month %in% c(4:7)) %>% 
  tidyr::drop_na() %>% 
  ggplot2::ggplot(.,aes(x=recruit_est, y=GSI)) + 
  geom_point(color = 'black') +
  xlab('Recruitment estimate') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index (Apr:Jul)') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()


  
```


Should I add a lag? If so, how much? 


```{r}

# recruitment index across all years 
dplyr::full_join(recruit, gsi.m %>%
                     group_by(year) %>% 
                     filter(month %in% c(3:8)) %>% 
                     summarise(m.gsi = mean(GSI)), 
                   by = join_by(year)) %>%
  ggplot2::ggplot(., aes(x=recruit_est, y=m.gsi)) + 
  geom_point(color = 'black') +
  geom_hline(yintercept = 0, lty = 2)+
  labs(title = 'All years') +
  xlab('Recruitment estimate') +
  ylab('Gulf stream position anomaly') +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()


# recruitment index across just the years that Paul recommended + el nino 1998  

tt = dplyr::full_join(recruit, gsi.m %>%
                     group_by(year) %>% 
                     filter(month %in% c(3:8)) %>% 
                     summarise(m.gsi = mean(GSI)), 
                   by = join_by(year)) 
  ggplot2::ggplot(tt %>% filter(year>1997), aes(x=recruit_est, y=m.gsi)) + 
  geom_point(color = 'black') +
  geom_hline(yintercept = 0, lty = 2)+
  labs(title = '1998:2020, No outliers') +
  xlab('Recruitment estimate') +
  ylab('Gulf stream position anomaly') +
  geom_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE,
              aes(colour = 'Linear')) + 
  geom_smooth(method = "lm", formula = y ~ x + I(x^2),
              size = 1, se = FALSE, aes(colour = 'Quadratic')) + 
  geom_smooth(method = "loess", formula = y ~ x, 
              size = 1, se = FALSE, aes(colour = 'Loess')) + 
  geom_smooth(method = "gam", formula = y ~  s(x), 
              size = 1, se = FALSE, aes(colour = 'Gam')) + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3),
              size = 1, se = FALSE, aes(colour = 'Gam2')) +
  scale_color_manual(name='Model',
                     breaks=c('Linear', 'Quadratic', 'Loess', 'Gam', 'Gam2'),
                     values=c('Linear'='black', 'Quadratic'='blue', 
                              'Loess'='red', 'Gam' = 'green',
                              'Gam2' = 'purple')) +
  theme_bw()






```


```{r}



ggplot(data = df, # add the data
       aes(x = year, y = m.gsi, # set x, y coordinates
           color = pos)) +    # color by GS position
  geom_boxplot() +
  facet_grid(~pos) + # create panes base on GS position
  ecodata::theme_facet()



# this gives main effects AND interactions
pos_aov <- aov(recruit_est ~ year * pos,
              data = df)

# look at effects and interactions
# summary(pos_aov)
tidy_pos_aov <- broom::tidy(pos_aov)
tidy_pos_aov
# write.csv(tidy_pos_aov, 'gs_pos_aov.csv')



ggplot(data = df, 
             aes(x = recruit_est/1000, y = m.gsi, fill = pos,  
                 group = year)) +
   geom_bar(color = "black", stat = "identity", 
            position = position_dodge2(preserve = "single"), width = 20) +
  theme_bw() +
  labs(title = 'All years',
       x = "\nRecruitment estimate (x1000)", 
       y = "Gulf stream position anomaly\n")

ggplot(data = df %>% filter(year <= 1999), 
       aes(x = recruit_est/1000, y = m.gsi, fill = pos,  
           group = year)) +
  geom_bar(color = "black", stat = "identity", 
           position = position_dodge2(preserve = "single"), width = 40) +
  theme_bw() +
  labs(title = '1971:1999',
       x = "\nRecruitment estimate (x1000)", 
       y = "Gulf stream position anomaly\n")

ggplot(data = df %>% filter(year >1999), 
       aes(x = recruit_est/1000, y = m.gsi, fill = pos,  
           group = year)) +
  geom_bar(color = "black", stat = "identity", 
           position = position_dodge2(preserve = "single"), width = 40) +
  theme_bw() +
  labs(title = '2000:2020',
       x = "\nRecruitment estimate (x1000)", 
       y = "Gulf stream position anomaly\n")



```



#### Food availablity 

Larval tilefish eat zooplankton, likely calanus. *Calanus finmarchicus* are a
copepod (crustacean) with a one-year life cycle and are an important food 
source for many commercially important species. Calanus spp. are lipid rich, 
herbivorous species that eat phytoplankton, diatoms in particular 
(Hobbs et al. 2020).

Diatoms are often represented as microplankton (>20 µm), but many species are of
the nanoplankton size class (2-20 µm), and a smaller few may even overlap 
with picoplanton size class (<2 µm).


##### Calanus

Calanus is not as common in MAB, need to figure out dominant zooplankton in 
MAB. 

```{r, c16}
# Calanus
calanus <- ecodata::calanus_stage %>% filter(Time %in% c(1998:2021))%>% 
    rename_all(., .funs = tolower) %>% 
   mutate(year = time)


ggplot() +
  geom_line(data = calanus %>% filter(epu == 'GB', 
                                  var == 'adt-Spring'), 
       aes(x = year , y = value, col = epu), lwd = 1) + 
  geom_line(data = calanus %>% filter(epu == 'MAB', 
                                  var == 'adt-Spring'), 
       aes(x = year , y = value, col = epu), lwd = 1) +
  labs(color = c('EPU')) +
  theme_minimal()

```


Georges Bank

```{r, c17}
# GB
c5.gb <- calanus %>% filter(epu == 'GB', var == 'c5-Spring')
adult.gb <- calanus %>% filter(epu == 'GB', var == 'adt-Spring' )

df.c5 <- dplyr::full_join(recruit, c5.gb, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()
df.adt <- dplyr::full_join(recruit, adult.gb, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()

# Regression
ggscatter(df.c5, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus c5 spring (No. per 100m^-3)",
           title = 'c5')  

ggscatter(df.adt, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus adult spring (No. per 100m^-3)",
           title = 'Adult')  
# GLM 
eqn <- as.formula(paste('recruit_est ~', paste(colnames(df.c5)[1], 
                                               collapse = " + ")))

mod0 <- glm(recruit_est ~ 1, 
            data = df.c5, 
            family = "poisson")

mod1 <- glm(eqn, 
            data = df.c5, 
            family = "poisson")

summary(mod0)
summary(mod1)
AIC(mod0, mod1) %>% dplyr::arrange(AIC)

null_prediction <- exp(predict(mod0))
mod_prediction <- exp(predict(mod1))


plot(df.c5$year, df.c5$recruit_est, type = 'l')
lines(df.c5$year, null_prediction, col = "gray")
lines(df.c5$year, mod_prediction, col = "red")

```



Mid-atlantic 

```{r, c18}
# Mid-Atlantic Bight
c5.mab <- calanus %>% filter(epu == 'MAB', var == 'c5-Spring')
adult.mab <- calanus %>% filter(epu == 'MAB', var == 'adt-Spring' )

df.c5 <- dplyr::full_join(recruit, c5.mab, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()
df.adt <- dplyr::full_join(recruit, adult.mab, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, value) %>%
  tidyr::drop_na()

# Regression
ggscatter(df.c5, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus c5 spring (No. per 100m^-3)",
           title = 'c5')  

ggscatter(df.adt, x = 'recruit_est', y = 'value', 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Recruitment estimate", 
          ylab = "Calanus adult spring (No. per 100m^-3)",
           title = 'Adult')  



```




##### Microplankton


##### Microplankton maps {.tabset}

```{r, micromaps, results='asis'}

# microplankton

```

##### Microplankton analysis

```{r}
# Microplankton
```




##### Chlorophyll-A
```{r}

# CHL-A

```

##### Maps (Chlorophyll-A) {.tabset}
```{r, chlmaps, results='asis'}

# CHL-A

```

##### Chlorophyll-A analysis

```{r, c21}

# CHL-A

```




##### SST Fronts

```{r}
# SST fronts

```

##### Maps (SST fronts) {.tabset}
```{r, sst front maps, results='asis'}


```

##### SST Fronts analysis

```{r}


```






References: 

Joyce, Terrence M, Young-Oh Kwon, Hyodae Seo, and Caroline C Ummenhofer. 2019.
“Meridional Gulf Stream Shifts Can Influence Wintertime Variability in the North 
Atlantic Storm Track and Greenland Blocking.” 
Geophysical Research Letters 46 (3): 1702–8. https://doi.org/10.1029/2018GL081087.

Hobbs, L., Banas, N. S., Cottier, F. R., Berge, J., & Daase, M. (2020). Eat or 
sleep: availability of winter prey explains mid-winter and spring activity in
an Arctic Calanus population. Frontiers in Marine Science, 7, 541564.



Pérez-Hernández, M. Dolores, and Terrence M. Joyce. 2014. 
“Two Modes of Gulf Stream Variability Revealed in the Last 
Two Decades of Satellite Altimeter Data.” 
Journal of Physical Oceanography 44 (1): 149–63. https://doi.org/10.1175/JPO-D-13-0136.1.













