---
title: "Environmental indicator analysis-Tilefish"
output:  
  html_document:
    code_fold: hide
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, wrangling}

# Loading packages and data
library(ggplot2)
library(ecodata)
library(lubridate)
library(dplyr)
library(stringr)
library(marmap) # bathymetry
library(RColorBrewer)
library(ggnewscale)
library(sf)
library(cowplot)
library(tidyverse)
library(ggpubr)
library(sf)
library(ggdist)
library(ggpubr)
library(wesanderson)
library(raster)
#library(glmTMB)
library(ggpmisc)
library(mgcViz)
library(gratia)

# CPUE data (no env covariates)
gt_data_model_cpue <- read.csv(here::here('data/catch_data/gt_data_model_cpue.csv'))
#names(gt_data_model_cpue) <- tolower(names(gt_data_model_cpue))
# Add in column with cpue 
# note: Paul indicated to use small mesh
gt_data_model_cpue <- gt_data_model_cpue %>% 
       rename_all(., .funs = tolower) %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM',mesh_size >= 5.6 ~ 'LG',
                                                      TRUE ~ 'NA')) %>%
  mutate(cpue_hr = sum_gt_catch/effort_dur)

# Catch data: 
sfobs <-readRDS(here::here('data/catch_data/gold_tile_sf_ob_v1_temp_price.rds'))

sfob.env <- sfobs %>% 
  mutate(mesh_bin = case_when(mesh_size <= 5.6 ~ 'SM', mesh_size >= 5.6 ~ 'LG',
                              TRUE ~ 'NA'),
         cpue_hr = SUM_GT_CATCH/effort_dur) %>% 
  filter(YEAR %in% c(1998:2022) & mesh_bin == 'SM') %>%
  dplyr::select(DATE, YEAR, MONTH, YDAY,trip_id,hull_num, area, effort_dur, 
         SUM_GT_CATCH, cpue_hr, mesh_size, mesh_bin, depth, start_lat, start_lon,
         bottomT, bottomT_avg, MIN_TEMP_C, MEAN_TEMP_C, MAX_TEMP_C,
         TEMP_VARIANCE, TEMP_DEVIATION, MEAN_DPTH_M, tri, sed) %>% 
  mutate(YEAR = as.integer(YEAR)) %>% 
  rename_all(., .funs = tolower)


areas <- sort(unique(sfob.env$area))
catch.tally.ann <- sfob.env %>% # aggregate by year
  group_by(year) %>% 
  summarise(ttl_sum = sum(sum_gt_catch))

# Length data from observer program 
lengths <- read.csv(here::here('data/catch_data/gt_data_length_andy.csv')) 
names(lengths) <- tolower(names(lengths))

# Recruitment estimates from 2021 report
recruit <- read.csv(here::here('data/assessment_data/tilefish_rec_estimate_2021.csv'))

# Merge SF/Obs catch data with recruit estimates:
catch_recruit <- cbind(recruit %>% filter(year %in% c(1998:2020)),
                       catch.tally.ann %>%
                         filter(year %in% c(1998:2020)) %>%
                         dplyr::select(ttl_sum))

# loading in shape files for maps
US.areas <- st_read(here::here('shapefiles/USA.shp'), quiet = TRUE)
canada.areas <- st_read(here::here('shapefiles/Canada.shp'), quiet = TRUE)
bts_strata <- st_read(here::here('shapefiles/NES_BOTTOM_TRAWL_STRATA.shp'),
                      quiet = TRUE)
# plot(bts_strata) # to see all bottom trawl strata

gtf_strata <- bts_strata %>% 
  filter(STRATUMA %in% c('01030', '01040', '01070', '01080', '01110', '01120', 
                         '01140', '01150', '01670', '01680', '01710', '01720', 
                         '01750', '01760')) # select just the gtf strata
# plot(gtf_strata)
bathy <- marmap::getNOAA.bathy(-81,-58, 27, 46)
bathy = fortify.bathy(bathy)

```

### Recruitment estimate

Figure 3. Age-1 recruitment estimate from the 2021 tilefish assessment across 
all years

```{r, c4}

ggplot(recruit, aes(x = factor(year), y = recruit_est, group = 1))+
   geom_rect(aes(xmin = '2007', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'red', alpha = 0.02) +
  geom_rect(aes(xmin = '2000', xmax = '2022', ymin = -Inf, ymax = Inf), 
            fill = 'lightblue', alpha = 0.05) +
   geom_vline(xintercept = c('1993','1999', '2005', '2013'), lty = 2) +
  geom_line(color = 'black', size = 1.5) +
  annotate("text", label = "*",
    x = 26, y = 14000, size = 8, colour = "red" )+
  xlab('Year') + 
  ylab('Total sum tilefish catch') + 
  # facet_wrap(~month)+
  theme(axis.text.x = element_text(color = 'black',
                                   size = 12, angle = 45, vjust = 1, hjust=1)) +
  ecodata::theme_facet()

### Thought: Should we isolate years associated w/strong year classes (or bad)
###           for correlations and analyses? 
```


##### SST

```{r, c9}
# SST
sst<-read.csv(here::here('data/sst/sst_ts_gtf_strata.csv'))
```


```{r}

# SST with season lag
sst.lag <- sst %>%
  arrange(year, month) %>%
  mutate(sst_lag3 = lag(weighted_mean_sst,3),
         sst_lag6 = lag(weighted_mean_sst,6),
         sst_lag9 = lag(weighted_mean_sst,9),
         sst_lag12 = lag(weighted_mean_sst,12))
         
# Join with recruit estimate
df <- dplyr::full_join(recruit, sst.lag %>%
                   group_by(year) %>% 
                   filter(year %in% c(2000:2020)),
dplyr::select(year, month, recruit.est, mean_sst, weighted_mean_sst, sst_lag3, sst_lag6, sst_lag9, sst_lag12),
                 by = join_by(year))
sst.rec <- df[-c(1:29),] #removes year < 2000 (when SST data begins)

#by season
sst_winter <- filter(sst.rec, month %in% c(1,2,3))
sst_spring <- filter(sst.rec, month %in% c(4,5,6))
sst_summer <- filter(sst.rec, month %in% c(7,8,9))
sst_fall <- filter(sst.rec, month %in% c(10,11,12))

```

Here we have SST in recruitment year - by month (updated 12/19)

```{r, nolagsst}
df = sst.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))


df %>% group_by(year, season, strong_year_class) %>%
  summarize(sst = mean(weighted_mean_sst),
            r = mean(recruit_est)) %>%
  ggscatter(x = "sst", y = "r",
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean Sea Surface Temperature", ylab = "Recruitment Estimate",
            title="Recruitment year - no lag") +
  facet_wrap(~season, scales = 'free')


```


Here we have SST lagged by 1 year - by season (updated 12/19)

```{r, laggedsst}
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sst = mean(sst_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sst", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean sea surface temperature", ylab = "Recruitment Estimate",
            title="Lag one year")+
  facet_wrap(~season, scales = 'free')
```

Seasonal lag

```{r}
## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sst = mean(sst_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sst", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean sea surface temperature", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sst = mean(sst_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sst", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean sea surface temperature", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sst = mean(sst_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sst", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean sea surface temperature", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sst = mean(sst_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sst", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean sea surface temperature", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")
```


##### Bottom Temperature

##### Bottom T analysis

The following figures compare in-situ bottom temperature from the study-fleet
data set to the recruitment estimate. 

+ Note here temperatures are averaged across all depths > 50 for each month.


```{r}
bt <- read.csv(here::here('data/bottom_temp/bt_ts_gtf_1970_2023.csv'))

# BT with year lag
bt.lag <- bt %>%
  filter(year %in% c(1997:2020)) %>%
  mutate(bt_lag3 = lag(weighted_mean_bt,3),
         bt_lag6 = lag(weighted_mean_bt,6),
         bt_lag9 = lag(weighted_mean_bt,9),
         bt_lag12 = lag(weighted_mean_bt,12))
         
# Join with recruit estimate
bt.rec <- dplyr::full_join(recruit, bt.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, weighted_mean_bt, bt_lag3, bt_lag6, bt_lag9, bt_lag12) %>%
  tidyr::drop_na()


#by season
bt_winter <- filter(bt.rec, month %in% c(1,2,3))
bt_spring <- filter(bt.rec, month %in% c(4,5,6))
bt_summer <- filter(bt.rec, month %in% c(7,8,9))
bt_fall <- filter(bt.rec, month %in% c(10,11,12))

```

Here is bottom temp and recruitment not lagged 

```{r, nolagbt}

df = bt.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))


df %>% group_by(year, season, strong_year_class) %>% 
  summarize(bt = mean(weighted_mean_bt),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean bottom temp", ylab = "Recruitment Estimate",
            title="Recruitment year - no lag")+
  facet_wrap(~season, scales = 'free')



```



Here we have BT lagged by 1 year - by month (updated 12/19)


```{r, lagbt}

df %>% group_by(year, season, strong_year_class) %>% 
  summarize(bt = mean(bt_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean bottom temp", ylab = "Recruitment Estimate",
            title="Lag one year")+
  facet_wrap(~season, scales = 'free')

```

Seasonal lag

```{r}
## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(bt = mean(bt_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean Bottom Temperature", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(bt = mean(bt_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean Bottom Temperature", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)
df %>% group_by(year, strong_year_class) %>% 
  summarize(bt = mean(bt_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean Bottom Temperature", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)
df %>% group_by(year, strong_year_class) %>% 
  summarize(bt = mean(bt_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "bt", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean Bottom Temperature", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")

```


##### Salinity

Here we explore salinity from the GLORYS reanalysis model at three different 
depths

 + 78 meters (relevant to larvae, recruits)
 + 92 meters (relevant recruits, juveniles)
 + 110 meters (relevant to juveniles, adults)


##### Salinity analysis


78 Meters 

```{r, 78m}
# Read in data at 78 m depth
sal78 <- read.csv(here::here('data/salinity/sal_78m_monthly_ts_gtf.csv'))

# sal78m with year lag
sal78.lag <- sal78 %>%
  filter(year %in% c(1997:2020)) %>%
  mutate(sal78_lag3 = lag(weighted_mean_sal_78m,3),
         sal78_lag6 = lag(weighted_mean_sal_78m,6),
         sal78_lag9 = lag(weighted_mean_sal_78m,9),
         sal78_lag12 = lag(weighted_mean_sal_78m,12), 
         year = as.numeric(year))
         
         
# Join with recruit estimate
sal78.rec <- dplyr::full_join(recruit, sal78.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, weighted_mean_sal_78m, sal78_lag3, sal78_lag6, sal78_lag9, sal78_lag12) %>%
  tidyr::drop_na()

```

No lag, 1 year lag, seasonal lags

```{r, nolagsal}

df = sal78.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal78 = mean(weighted_mean_sal_78m),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="Recruit year - no lag 78m")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal78 = mean(sal78_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="Lag one year 78m")+
  facet_wrap(~season, scales = 'free')

#### Computing correlations
# win.cor = df %>% filter(season == 'winter') 
# print(paste0('correlation coeff: ', round(cor(win.cor[,3], win.cor[,5]), 4)))
# fal.cor = df %>% filter(season == 'fall') 
# print(paste0('correlation coeff: ', round(cor(fal.cor[,3], fal.cor[,4]), 4)))

## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal78 = mean(sal78_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal78 = mean(sal78_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal78 = mean(sal78_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal78 = mean(sal78_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal78", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (78m)", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")

```




92 Meters 

```{r, 92m}
# Read in data at 92 meter depth 

sal92 <- read.csv(here::here('data/salinity/sal_92m_monthly_ts_gtf.csv'))

sal92.lag <- sal92 %>%
  filter(year %in% c(1997:2020)) %>%
  mutate(sal92_lag3 = lag(weighted_mean_sal_92m,3),
         sal92_lag6 = lag(weighted_mean_sal_92m,6),
         sal92_lag9 = lag(weighted_mean_sal_92m,9),
         sal92_lag12 = lag(weighted_mean_sal_92m,12), 
         year = as.numeric(year))
         
         
# Join with recruit estimate
sal92.rec <- dplyr::full_join(recruit, sal92.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, weighted_mean_sal_92m, sal92_lag3, sal92_lag6, sal92_lag9, sal92_lag12) %>%
  tidyr::drop_na()

```

No lag, 1 year lag, and seasonal lag plots

```{r, nolagsal2}

df = sal92.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal92 = mean(weighted_mean_sal_92m),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="Recruit year - no lag 92m")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal92 = mean(sal92_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="Lag one year 92m")+
  facet_wrap(~season, scales = 'free')

#### Computing correlations
# win.cor = df %>% filter(season == 'winter') 
# print(paste0('correlation coeff: ', round(cor(win.cor[,3], win.cor[,5]), 4)))
# fal.cor = df %>% filter(season == 'fall') 
# print(paste0('correlation coeff: ', round(cor(fal.cor[,3], fal.cor[,4]), 4)))

## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal92 = mean(sal92_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal92 = mean(sal92_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal92 = mean(sal92_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal92 = mean(sal92_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal92", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (92m)", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")

#### Computing correlations
# win.cor = df %>% filter(season == 'winter') 
# print(paste0('correlation coeff: ', round(cor(win.cor[,3], win.cor[,5]), 4)))
# fal.cor = df %>% filter(season == 'fall') 
# print(paste0('correlation coeff: ', round(cor(fal.cor[,3], fal.cor[,4]), 4)))


```







110 Meters 

```{r, 110m}
# Wrangling the data
sal110 <- read.csv(here::here('data/salinity/sal_110m_monthly_ts_gtf.csv'))

#sal110 with lags
sal110.lag <- sal110 %>%
  filter(year %in% c(1997:2020)) %>%
  mutate(sal110_lag3 = lag(weighted_mean_sal_110m,3),
         sal110_lag6 = lag(weighted_mean_sal_110m,6),
         sal110_lag9 = lag(weighted_mean_sal_110m,9),
         sal110_lag12 = lag(weighted_mean_sal_110m,12), 
         year = as.numeric(year))
         
         
# Join with recruit estimate
sal110.rec <- dplyr::full_join(recruit, sal110.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, weighted_mean_sal_110m, sal110_lag3, sal110_lag6, sal110_lag9, sal110_lag12) %>%
  tidyr::drop_na()


```

No lag, 1 year lag, and seasonal lag plots

```{r, sal110reg}

df = sal110.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal110 = mean(weighted_mean_sal_110m),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="Recruit year - no lag 110m")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(sal110 = mean(sal110_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="Lag one year 110m")+
  facet_wrap(~season, scales = 'free')

#### Computing correlations
# win.cor = df %>% filter(season == 'winter') 
# print(paste0('correlation coeff: ', round(cor(win.cor[,3], win.cor[,5]), 4)))
# fal.cor = df %>% filter(season == 'fall') 
# print(paste0('correlation coeff: ', round(cor(fal.cor[,3], fal.cor[,4]), 4)))

## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal110 = mean(sal110_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal110 = mean(sal110_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal110 = mean(sal110_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(sal110 = mean(sal110_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "sal110", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean salinity (110m)", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")

```


##### Cold pool analysis
Recruit with no lag has a significant negative relationship with increased
cold pool extent. 


Excerpt from New York Bight Fish, Fisheries, and Sand Features: Data Review
Volume 1: Literature Synthesis and Gap Analysis: 

Natural events are documented for Scup (Morse 1978; Smith and Norcross 1968), Bluefish (Slater et al. 2007), and Northern Puffer (Wicklund 1970) in the MAB and NYB as a result of upwelling of cold bottom water, and for Golden Tilefish on the OCS from eastward flow of the Cold Pool (Fisher et al. 2014). Such events are
difficult to detect and quantify because they are episodic, because they could happen in winter when there are few observers, or because they leave dead animals on the bottom instead of floating (Fisher et al.2014) so that they are only detected incidentally by trawler (e.g., Woodhead (1964)).

```{r}
library(ecodata)
# Pull cold pool data
cp.ex <- ecodata::cold_pool
# Selecting just the extent
cp <- cp.ex %>% filter(Time %in% c(1994:2020), 
              Var == 'extent_index') %>%
  mutate(ext = Value,
         ext_lag1 = lag(ext,1), 
         year = as.numeric(Time))
## NOTE: I added more lags (up to 7), but they only got worse - may make more 
## sense to do more lags with cpue data with older fish


# Join with recruit estimate 
cp.rec <- dplyr::full_join(recruit, cp, by = join_by(year)) %>%
  dplyr::select(year, recruit_est, ext, ext_lag1) %>% 
  filter(year %in% c(1994:2020))

cp.rec %>% 
  ggscatter(x = 'ext', y = 'recruit_est', 
            add = "reg.line", conf.int = TRUE, 
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Cold pool extent", ylab = "Recruitment Estimate",
            title="Recruit year")

cp.rec %>% 
  ggscatter(x = 'ext_lag1', y = 'recruit_est', 
            add = "reg.line", conf.int = TRUE, 
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Cold pool extent", ylab = "Recruitment Estimate",
            title="Lag 1 year")

```



#### Currents

Cross-shelf processes may influence the retention or displacement of tilefish 
during early life history stages. These are explored below.

##### Shelf water volume

Shelf water volume: A measure of the volume of water bounded inshore of the
shelf-slope front. In this analysis, shelf water is defined as all water having 
salinity <34 psu. The position of the shelf-slope front varies inter-annually
with the higher shelf water values indicating the front being pushed further 
towards the shelf break.

high shv: front pushed towards sbf
low shv: front pushed inshore (more slope water on shelf)

Hypothesis: Higher recruitment success correlated with years of higher shelf 
water volume in spring/summer. These months months may be particularly 
important as that is when spawning is occurring and the position of the sbf
may influence the position of larvae (away from spawning grounds).

Additional variables in this dataset are shelf water temperature and salinity 
which may also be indicative of habitat conditions. 


```{r, s.w.volume}

# Shelf water volume
shlfvol <- read.csv(here::here('data/shelf_water_volume/ShelfWaterVolume_BSB_update.csv'))

# wrangling date info, converting doy to date and month 
yrs <- as.vector(nrow(shlfvol))
shlfvol$Year <- as.character(shlfvol$Year)
for (i in 1:nrow(shlfvol)){
  yrs[i] <- strsplit(shlfvol$Year, ".", fixed = TRUE)[[i]][1]
}
shlfvol$year <- yrs

shlfvol <- shlfvol %>% mutate(date_= as.Date(Year.Day-1, 
                                             origin=paste0(year, "-01-01")), 
                              month= strftime(date_, "%m"), 
                              day=strftime(date_,"%d"), 
                              year = as.integer(year),
                              month = as.numeric(month))  


# Create shw vol by month w/lag
df.lag = shlfvol %>%
  group_by(year,month) %>% 
  summarise(mean_t = mean(ShW.T),
            mean_s = mean(ShW.S),
            mean_v = mean(ShW.Vol)) %>% 
  mutate(mean_t_lag2 = lag(mean_t,2),
         mean_t_lag3 = lag(mean_t,3), 
         mean_t_lag6 = lag(mean_t,6),
         mean_s_lag2 = lag(mean_s,2),
         mean_s_lag3 = lag(mean_s,3), 
         mean_s_lag6 = lag(mean_s,6),
         mean_v_lag2 = lag(mean_v,2),
         mean_v_lag3 = lag(mean_v,3), 
         mean_v_lag6 = lag(mean_v,6))
# Join in-situ bottom temps w/assessment recruitment estimate
df.join.shlf = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_t, mean_s, mean_v, 
                mean_t_lag2, mean_s_lag2, mean_v_lag2,
                mean_t_lag3, mean_s_lag3, mean_v_lag3) %>%
  tidyr::drop_na()
  

```

##### Shelf Water Volume Analysis

```{r}
df.join.shlf = dplyr::full_join(recruit, df.lag, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, mean_v) %>%
  tidyr::drop_na() #removing lags to visualize all months
 
df.join.shlf = df.join.shlf %>%
  filter(year %in% c(1998:2020)) %>%
  mutate(v_lag3 = lag(mean_v,3),
         v_lag6 = lag(mean_v,6),
         v_lag9 = lag(mean_v,9),
         v_lag12 = lag(mean_v,12)) #creating seasonal and one year lags

df = df.join.shlf %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))


# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(mean_v),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="Recruit year - no lag")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(v_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="Recruit year - 1 year lag")+
  facet_wrap(~season, scales = 'free')

## 3 month lag (winter)

df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(v_lag3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(v_lag6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(v_lag9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, season, strong_year_class) %>% 
  summarize(v = mean(v_lag12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "v", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Mean shelf water volume", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")
```


##### Gulf-stream index

Gulf stream index was calculated based on method presented by 
Pérez-Hernández and Joyce (2014). The gulf stream index (GSI) is a measure of 
the degrees latitude above the average Gulf Stream position based on ocean
temperature at 200m (15 C) depth between 55W to 75W.

Positive values indicate a the mean position of the GS is more Northernly, 
whereas negative values indicate a more Southernly position. 

```{r}
gsi.m <- read.csv(here::here('data/gulf_stream_index/mm_gsi_1954_2022_chen.csv'))


## --- This joins recruit estimate data to GSI data and 
## --- calculates some summary stats(mean, sd, min, max)
df <- dplyr::full_join(recruit, gsi.m %>%
                   group_by(year) %>% 
                   filter(month %in% c(3:8)) %>% 
                   summarise(m.gsi = mean(GSI),
                             sd.gsi = sd(GSI),
                             max.gsi = max(GSI), 
                             min.gsi = min(GSI)), 
                 by = join_by(year)) %>% 
  mutate(pos = ifelse(m.gsi > 0, 'Northerly', 'Southerly'), 
         n.pos = ifelse(m.gsi > 0, 1, 0))
df <- df[-c(51:69),] # this removes rows w/years that don't match 
# (could/should do this before hand in more standardized way - will do later)

# Plot the time series (mean GSI for the months of interest)
# -- Note: here I am looking just at March through August assuming these
# months are the most important to recently spawned individuals 
ggplot(data = df, 
       aes(x = year, y = m.gsi)) +
  geom_line(lwd = 1) +
  geom_hline(yintercept = 0, lty = 2) +
  labs(title = 'Mean GS position anomaly March:August',
       x = 'Year', 
       y = "Gulf stream position anomaly\n") +
theme_bw() 

# Looking across all months 
dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, GSI) %>%
  #filter(month %in% c(3:9)) %>%  
  filter(year>1997) %>% 
  tidyr::drop_na() %>% 
ggplot2::ggplot(.,aes(x=recruit_est, y=GSI)) + 
  geom_point(color = 'black') +
  facet_wrap(~month)+
  xlab('Recruitment estimate') +
  ylab('Gulf stream position anomaly') +
  labs(title = 'Gulf stream index by month') +
  geom_hline(yintercept = 0, lty = 2) +
  theme_bw()
 
```

##### Gulf Stream Index Analysis

```{r}
df.gsi = dplyr::full_join(recruit, gsi.m, by = join_by(year)) %>%
  dplyr::select(year, month, recruit_est, GSI) %>%
  mutate(gsi_3 = lag(GSI,3),
         gsi_6 = lag(GSI,6),
         gsi_9 = lag(GSI,9),
        gsi_12 = lag(GSI,12)) %>%
  filter(year>1997) %>% 
  tidyr::drop_na() #GSI by month and with year lag

df = df.gsi %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(gsi = mean(GSI),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream position anomaly", ylab = "Recruitment Estimate",
            title="Recruit year - no lag")+
  facet_wrap(~season, scales = 'free')

# Plot with 1 year lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(gsi = mean(gsi_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream Position Anomaly", ylab = "Recruitment Estimate",
            title="Lag one year")+
  facet_wrap(~season, scales = 'free')

## Seasonal lags

# 3 month lag - Winter
df %>% group_by(year, strong_year_class) %>% 
  summarize(gsi = mean(gsi_3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream Position Anomaly", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

# 6 month lag - Winter
df %>% group_by(year, strong_year_class) %>% 
  summarize(gsi = mean(gsi_6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream Position Anomaly", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

# 9 month lag - Winter
df %>% group_by(year, strong_year_class) %>% 
  summarize(gsi = mean(gsi_9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream Position Anomaly", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

# 12 month lag - Winter
df %>% group_by(year, strong_year_class) %>% 
  summarize(gsi = mean(gsi_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "gsi", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Gulf Stream Position Anomaly", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")
```

#### Food availablity 

Larval tilefish eat zooplankton, likely calanus. *Calanus finmarchicus* are a
copepod (crustacean) with a one-year life cycle and are an important food 
source for many commercially important species. Calanus spp. are lipid rich, 
herbivorous species that eat phytoplankton, diatoms in particular 
(Hobbs et al. 2020).

Diatoms are often represented as microplankton (>20 µm), but many species are of
the nanoplankton size class (2-20 µm), and a smaller few may even overlap 
with picoplanton size class (<2 µm).


##### Microplankton
```{r}
micro<-read.csv(here::here('data/phyto_size_class/microplankton_ts_gtf_strata.csv'))

# micro with year lag
micro.lag <- micro %>%
  arrange(year, month) %>%
  mutate(micro_3 = lag(weighted_mean_micro,3),
         micro_6 = lag(weighted_mean_micro,6),
         micro_9 = lag(weighted_mean_micro,9),
         micro_12 = lag(weighted_mean_micro,12))
         
# Join with recruit estimate
micro.rec <- dplyr::full_join(recruit, micro.lag %>%
                   group_by(year) %>% 
                   filter(year %in% c(1998:2020)),
dplyr::select(year, month, mean_cpue, mean_micro, weighted_mean_micro, micro_3, micro_6, micro_9, micro_12),
                 by = join_by(year))
micro.rec <- micro.rec[-c(1:27),] #removes year < 1998 (starting year for analysis)
```


##### Microplankton maps {.tabset}

```{r, micromaps, results='asis'}

# microplankton

```

##### Microplankton analysis

```{r}
df = micro.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(micro = mean(weighted_mean_micro),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean microplankton", ylab = "Recruitment Estimate",
            title="Recruit year - no lag")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(micro = mean(micro_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted Mean Microplankton", ylab = "Recruitment Estimate",
            title="Lag one year")+
  facet_wrap(~season, scales = 'free')

```

Seasonal lags

```{r}
## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(micro = mean(micro_3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean microplankton", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(micro = mean(micro_6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean microplankton", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(micro = mean(micro_9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean microplankton", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(micro = mean(micro_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "micro", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean microplankton", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")
```



##### Chlorophyll-A
```{r}

# CHL-A
chl<-read.csv(here::here('data/chl/chl_ts_gtf_strata.csv'))

# chl with year lag
chl.lag <- chl %>%
  arrange(year, month) %>%
  mutate(chl_3 = lag(weighted_mean_chl,3),
         chl_6 = lag(weighted_mean_chl,6),
         chl_9 = lag(weighted_mean_chl,9),
         chl_12 = lag(weighted_mean_chl,12))
         
# Join with recruit estimate
chl.rec <- dplyr::full_join(recruit, chl.lag %>%
                   group_by(year) %>% 
                   filter(year %in% c(1998:2020)),
dplyr::select(year, month, mean_cpue, mean_chl, weighted_mean_chl, chl_3, chl_6, chl_9, chl_12),
                 by = join_by(year))
chl.rec <- chl.rec[-c(1:27),] #removes year < 1998 (starting year for analysis)
```

##### Maps (Chlorophyll-A) {.tabset}
```{r, chlmaps, results='asis'}

# CHL-A

```

##### Chlorophyll-A analysis

```{r, c21}
df = chl.rec %>% 
  mutate(season = case_when(month %in% c(1,2,3) ~ 'winter', 
                            month %in% c(4,5,6) ~ 'spring', 
                            month %in% c(7,8,9) ~ 'summer', 
                            month %in% c(10,11,12) ~ 'fall'),
         strong_year_class = case_when(year %in% c(1998,1999,2005,2013,2017) ~ 'yes',
                                  year %in% c(2000,2006,2014,2018) ~ 'year after',
                                  year %in% c(2001:2004,2007:2012,2015:2016,2019:2020) ~ 'no'))

# Plot without lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(chl = mean(weighted_mean_chl),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
             color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="Recruit year - no lag")+
  facet_wrap(~season, scales = 'free')

# Plot with lag
df %>% group_by(year, season, strong_year_class) %>% 
  summarize(chl = mean(chl_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
             color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="Lag one year")+
  facet_wrap(~season, scales = 'free')
```

Seasonal lags

```{r}
## 3 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(chl = mean(chl_3),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="3 month lag - Winter")

## 6 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(chl = mean(chl_6),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="6 month lag - Winter")

## 9 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(chl = mean(chl_9),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="9 month lag - Winter")

## 12 month lag (winter)

df %>% group_by(year, strong_year_class) %>% 
  summarize(chl = mean(chl_12),
            r = mean(recruit_est)) %>% 
  ggscatter(x = "chl", y = "r", 
            add = "reg.line", conf.int = TRUE, 
            color = "strong_year_class", palette = c("black", "green", "red"),
            cor.coef = TRUE, cor.coeff.args = list(method="pearson"),
            add.params=list(color="dodgerblue", fill="lightgray"),
            xlab = "Weighted mean chlorophyll-a", ylab = "Recruitment Estimate",
            title="12 month lag - Winter")
```




##### SST Fronts

```{r}
# SST fronts
fprob<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_3x3.csv')) #fprob across all individual strata
fprob.ind<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_indv_substrata_3x3.csv')) #mean for each of 14 substrata
fprob.ns<-read.csv(here::here('data/sst_fronts/fprob_seasonal_ts_gtf_n_v_s_substrata_3x3.csv')) #mean for substrata N and S of Hudson canyon

# fprob with year lag
fprob.lag <- fprob %>%
  mutate(fprob_lag1 = lag(weighted_mean_fprob,4))
         
# Join with recruit estimate
fprob.rec <- dplyr::full_join(recruit, fprob.lag %>%
                   group_by(year) %>% 
                   filter(year %in% c(1998:2020)),
dplyr::select(year, season, recruit_est, mean_fprob, weighted_mean_fprob, fprob_lag1),
                 by = join_by(year))
fprob.rec <- fprob.rec[-c(1:29),] #removes year < 2000 (starting year for fprob)
```

##### Maps (SST fronts) {.tabset}
```{r, sst front maps, results='asis'}


```

##### SST Fronts analysis

```{r}


```

**Fprob of each substrata**

```{r}
#join data
ind.rec <- dplyr::full_join(recruit, fprob.ind %>%
                   group_by(year) %>% 
                   filter(year %in% c(1998:2020)),
dplyr::select(year, season, recruit_est, mean_fprob, weighted_mean_fprob, fprob_lag1),
                 by = join_by(year))
ind.rec <- ind.rec[-c(1:29),]  #begin at year 2000



```

**Fprob N/S of Hudson Canyon**

```{r}
#join data
ns.rec <- dplyr::full_join(recruit, fprob.ns %>%
                   group_by(year) %>% 
                   filter(year %in% c(1998:2020)),
dplyr::select(year, season, recruit_est, mean_fprob, weighted_mean_fprob, fprob_lag1),
                 by = join_by(year))
ns.rec <- ns.rec[-c(1:29),]  #begin at year 2000


```

##### GAMs

```{r}
## habitat (sst/bt/sal?)
habitat <- dplyr::full_join(
  sst.rec, bt.rec, by = join_by(year, month, recruit_est)) %>%
  dplyr::select(year, month, recruit_est, weighted_mean_sst, weighted_mean_bt)


habitat.gam <- gam(recruit_est ~ s(weighted_mean_sst, k=10) + s(weighted_mean_bt, k=10), data=habitat, method="REML")
summary(habitat.gam)
plot(habitat.gam, seWithMean=TRUE)

##currents (shelfvol/gsi)
currents <- dplyr::full_join(
  df.join.shlf, df.gsi, by = join_by(year, month, recruit_est)) %>%
  dplyr::select(year,month, recruit_est, mean_v, GSI)
#no gsi <1998, no shlfvol 2020 >

currents.gam <- gam(recruit_est ~ s(mean_v, k=10) + s(GSI, k=10), data=currents, method="REML")
summary(currents.gam)
plot(currents.gam, seWithMean=TRUE)

## food availability (micro/chl-a/sstfronts)
food1<- dplyr::full_join(
  micro.rec, chl.rec, by = join_by(year, month, recruit_est)) %>%
  dplyr::select(year,month, recruit_est, weighted_mean_micro, weighted_mean_chl)

food<-read.csv(here::here('data/food_rec.csv')) #join with fprob and grouped by season

food.gam <- gam(recruit_est ~ s(weighted_mean_micro, k=10) + s(weighted_mean_chl, k=10) + s(weighted_mean_fprob, k=10), data=food, method="REML")
summary(food.gam)
plot(food.gam, seWithMean=TRUE)
```



References: 

Joyce, Terrence M, Young-Oh Kwon, Hyodae Seo, and Caroline C Ummenhofer. 2019.
“Meridional Gulf Stream Shifts Can Influence Wintertime Variability in the North 
Atlantic Storm Track and Greenland Blocking.” 
Geophysical Research Letters 46 (3): 1702–8. https://doi.org/10.1029/2018GL081087.

Hobbs, L., Banas, N. S., Cottier, F. R., Berge, J., & Daase, M. (2020). Eat or 
sleep: availability of winter prey explains mid-winter and spring activity in
an Arctic Calanus population. Frontiers in Marine Science, 7, 541564.



Pérez-Hernández, M. Dolores, and Terrence M. Joyce. 2014. 
“Two Modes of Gulf Stream Variability Revealed in the Last 
Two Decades of Satellite Altimeter Data.” 
Journal of Physical Oceanography 44 (1): 149–63. https://doi.org/10.1175/JPO-D-13-0136.1.













